---
title: "Validation of {reappraised} for continuous data"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

This validation uses the `reappraised` R package's internal functions directly (via `:::`), rather than copy-pasted code from the app. It checks whether `reappraised`'s p-value bounds (computed from rounded summary statistics) correctly contain the true p-value from simulated data.

reappraised version 0.6

```{r include=FALSE}

# formatting options
# set default chunk options
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

# disable scientific notation
options(scipen = 999)

```

# Dependencies

```{r}

library(tibble)
library(roundwork)
library(tidyr)
library(dplyr)
library(readr)
library(purrr)
library(furrr)
library(ggplot2)
library(scales)
library(knitr)
library(kableExtra)
library(parameters)
library(reappraised)

```

# Functions

```{r}

# --- Data generation ---

generate_data <- function(m1, m2, sd1, sd2, n1, n2) {
  data1 <- tibble(condition = "1",
                  score = rnorm(n = n1, mean = m1, sd = sd1))
  data2 <- tibble(condition = "2",
                  score = rnorm(n = n2, mean = m2, sd = sd2))
  bind_rows(data1, data2)
}

# --- Analysis: compute summary stats and true p ---

analyze <- function(data, m_digits, sd_digits, p_digits, analysis_rounding) {

  res_summary <- data |>
    group_by(condition) |>
    summarize(m = mean(score),
              sd = sd(score),
              n = n()) |>
    pivot_wider(names_from = condition,
                names_sep = "",
                values_from = c("m", "sd", "n"))

  res_ttest <-
    t.test(formula = score ~ condition,
           data = data,
           var.equal = TRUE,
           alternative = "two.sided") |>
    model_parameters() |>
    as_tibble() |>
    select(t, df = df_error, p)

  res <- bind_cols(res_summary, res_ttest)

  # Round summary stats as if they were reported
  round_fn <- switch(analysis_rounding,
    "half_up"   = roundwork::round_up,
    "half_down" = roundwork::round_down,
    "bankers"   = base::round,
    "trunc"     = roundwork::round_trunc
  )

  res <- res |>
    mutate(m1 = round_fn(m1, m_digits),
           m2 = round_fn(m2, m_digits),
           sd1 = round_fn(sd1, sd_digits),
           sd2 = round_fn(sd2, sd_digits),
           p_unrounded = p,
           p = round_fn(p, p_digits),
           p_operator = if_else(near(p, 0), "less_than", "equals"),
           p = if_else(near(p, 0), 1 * 10^-p_digits, p))

  return(res)
}

# --- Reconstruct bounds using reappraised package functions ---

reconstruct_reappraised <- function(results, m_digits, sd_digits, p_digits) {

  # Format p as character with correct decimal places (as it would appear in a paper)
  if (results$p_operator == "less_than") {
    p_char <- paste0("<", formatC(results$p, format = "f", digits = p_digits))
  } else {
    p_char <- formatC(results$p, format = "f", digits = p_digits)
  }

  # Build wide-format data as load_clean would produce
  dat <- tibble(
    study = "s1", var = "v1",
    m1 = results$m1, m2 = results$m2,
    s1 = results$sd1, s2 = results$sd2,
    n1 = results$n1, n2 = results$n2,
    p = p_char,
    p_num = results$p,
    dpm1 = reappraised:::dp(results$m1),
    dpm2 = reappraised:::dp(results$m2),
    dps1 = reappraised:::dp(results$sd1),
    dps2 = reappraised:::dp(results$sd2),
    dpp = as.numeric(p_digits),
    data_type = "pval_cont"
  )

  # Compute p-value from summary stats using reappraised's ANOVA method
  b <- reappraised:::lng_f(
    dat |> select(-c(p, p_num, data_type, dpp)),
    remvars = c("study", "var")
  )
  p_calc <- reappraised:::pval(b)

  # Merge and create p_dec.place (replicating comp_p logic)
  a1 <- left_join(dat, p_calc, by = c("study", "var")) |>
    mutate(
      p_dp = dpp,
      p.value = round(p.value, p_dp),
      diff_rep_calc_p = round(abs(p.value - p_num), p_dp)
    ) |>
    rename(p.value_calc = p.value, p_dec.place = p_dp)

  # Compute bounds using chk_pvals
  nm <- c("m1", "s1", "m2", "s2")
  nm_m <- c("m1", "m2")
  nm_s <- c("s1", "s2")
  nm_n <- c("n1", "n2")
  nm.m <- paste0("dp", nm_m)
  nm.s <- paste0("dp", nm_s)
  nm.sf <- paste0(nm, "_sf")

  am <- dat |> select(-c(p_num, data_type))

  result <- suppressWarnings(
    reappraised:::chk_pvals(
      dat = a1,
      am = am,
      nm = nm, nm_m = nm_m, nm_s = nm_s, nm_n = nm_n,
      nm.m = nm.m, nm.s = nm.s, nm.sf = nm.sf
    )
  )

  tibble(
    p_min_reappraised = result$p_any_min,
    p_max_reappraised = result$p_any_max,
    p_flag_reappraised = result$p_flag_any,
    p_inbounds_reappraised = (result$p_flag_any == "" | is.na(result$p_flag_any))
  )
}

# --- Naive p (ignoring rounding) ---

calculate_naive_p <- function(results) {
  diff_mean <- results$m1 - results$m2
  df_s <- results$n1 + results$n2 - 2
  sp <- sqrt(((results$n1 - 1) * results$sd1^2 + (results$n2 - 1) * results$sd2^2) / df_s)
  se_pooled <- sp * sqrt(1 / results$n1 + 1 / results$n2)
  t_val <- diff_mean / se_pooled
  p_val <- 2 * (1 - pt(abs(t_val), df = df_s))

  data.frame(naive_p = p_val)
}

```

# Run simulation

```{r}

if (file.exists("simulation_reappraised_cont.rds")) {
  simulation <- read_rds("simulation_reappraised_cont.rds")
} else {

  # simulation parameters
  experiment_parameters <- expand_grid(
    pop_m1 = 0,
    pop_m2 = c(0, 0.5, 2),
    pop_sd1 = 1,
    pop_sd2 = 1,
    pop_n1 = 25,
    pop_n2 = 25,
    m_digits = 2,
    sd_digits = 2,
    p_digits = 3,
    analysis_rounding = c("half_up", "half_down", "bankers", "trunc"),
    iteration = 1:1000
  )

  # set up parallelization
  plan(multisession, workers = future::availableCores())

  # set seed
  set.seed(42)

  # run simulation
  simulation <- experiment_parameters |>
    # generate datasets
    mutate(generated_data = future_pmap(
      .l = list(m1 = pop_m1, m2 = pop_m2, sd1 = pop_sd1, sd2 = pop_sd2, n1 = pop_n1, n2 = pop_n2),
      .f = generate_data,
      .progress = TRUE,
      .options = furrr_options(seed = TRUE)
    )) |>
    # analyze data to create the real p and summary stats
    mutate(results_original = future_pmap(
      .l = list(data = generated_data, m_digits = m_digits, sd_digits = sd_digits,
                p_digits = p_digits, analysis_rounding = analysis_rounding),
      .f = analyze,
      .progress = TRUE,
      .options = furrr_options(seed = TRUE)
    )) |>
    # reconstruct bounds using reappraised package functions
    mutate(results_reappraised = future_pmap(
      .l = list(results = results_original, m_digits = m_digits,
                sd_digits = sd_digits, p_digits = p_digits),
      .f = reconstruct_reappraised,
      .progress = TRUE,
      .options = furrr_options(seed = TRUE)
    )) |>
    unnest(results_reappraised) |>
    # naive p (ignoring rounding)
    mutate(results_naive = future_map(
      .x = results_original,
      .f = calculate_naive_p,
      .progress = TRUE,
      .options = furrr_options(seed = TRUE)
    )) |>
    unnest(results_naive)

  write_rds(simulation, "simulation_reappraised_cont.rds", compress = "gz")

}

```

```{r}

# extract reported p and p_operator for downstream use
simulation <- simulation |>
  mutate(p = map_dbl(results_original, ~ .x$p),
         p_operator = map_chr(results_original, ~ .x$p_operator),
         p_beyond_bounds_reappraised = case_when(
           p < p_min_reappraised ~ p - p_min_reappraised,
           p > p_max_reappraised ~ p_max_reappraised - p,
           TRUE ~ 0
         ),
         p_abs_beyond_bounds_reappraised = abs(p_beyond_bounds_reappraised),
         p_naive_p_diff = naive_p - p)

```

# Summarize results

## Coverage

All results:

```{r}

simulation |>
  summarize(
    n = n(),
    coverage_p_reappraised = mean(p_inbounds_reappraised)
  ) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

Excluding cases where the reported p was at the floor (p_operator == "less_than", i.e., p reported as "< 0.001"). These are expected to flag because the exact reported p (0.001) exceeds bounds when the true p is effectively 0:

```{r}

simulation |>
  filter(p_operator == "equals") |>
  summarize(
    n = n(),
    coverage_p_reappraised = mean(p_inbounds_reappraised)
  ) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

By factorial:

```{r}

simulation_coverage_by <- simulation |>
  group_by(pop_es = pop_m2, analysis_rounding, p_operator) |>
  summarize(
    n = n(),
    coverage_p_reappraised = mean(p_inbounds_reappraised)
  )

simulation_coverage_by |>
  kable(digits = 4) |>
  kable_classic(full_width = FALSE)

```

Only cells with imperfect coverage, excluding less_than cases:

```{r}

simulation_coverage_by |>
  filter(p_operator == "equals") |>
  filter(!near(coverage_p_reappraised, 1)) |>
  kable(digits = 4) |>
  kable_classic(full_width = FALSE)

```

## Visualizations

```{r fig.height=5, fig.width=7}

simulation |>
  ggplot(aes(x = p, y = naive_p, color = p_inbounds_reappraised)) +
  geom_abline(slope = 1, linetype = "dotted") +
  geom_linerange(aes(ymin = p_min_reappraised, ymax = p_max_reappraised), alpha = 0.4) +
  geom_point(alpha = 0.4) +
  scale_color_viridis_d(begin = 0.2, end = 0.5) +
  theme_linedraw() +
  facet_wrap(~ analysis_rounding) +
  ggtitle("Reappraised package: p-value bounds vs. naive p")

```

False positives only (true p outside bounds):

```{r fig.height=5, fig.width=7}

simulation |>
  filter(!p_inbounds_reappraised) |>
  ggplot(aes(x = p, y = naive_p, color = p_inbounds_reappraised)) +
  geom_abline(slope = 1, linetype = "dotted") +
  geom_linerange(aes(ymin = p_min_reappraised, ymax = p_max_reappraised), alpha = 0.6) +
  geom_point(alpha = 0.6) +
  theme_linedraw() +
  facet_wrap(~ analysis_rounding) +
  ggtitle("Reappraised package: false positives\n(true p outside computed bounds)")

```

## Bounds width

```{r}

simulation |>
  mutate(p_bounds_width = p_max_reappraised - p_min_reappraised,
         population_d = paste0("Population Cohen's d = ", pop_m2)) |>
  group_by(population_d, analysis_rounding) |>
  summarize(mean_p_bounds_width = mean(p_bounds_width)) |>
  kable(digits = 4) |>
  kable_classic(full_width = FALSE)

```

```{r fig.height=5, fig.width=7}

simulation |>
  mutate(p_bounds_width = p_max_reappraised - p_min_reappraised,
         population_d = paste0("Population Cohen's d = ", pop_m2)) |>
  ggplot(aes(p_bounds_width)) +
  geom_histogram(boundary = 0) +
  facet_grid(analysis_rounding ~ population_d) +
  theme_linedraw()

```

## Precision of naive p-values

```{r}

simulation |>
  summarise(
    p_bias_mean = mean(p_naive_p_diff),
    p_accuracy_rmse = sqrt(mean(p_naive_p_diff^2)),
    p_accuracy_abs_max = max(abs(p_naive_p_diff))
  ) |>
  mutate(across(everything(), ~ round(.x, 4))) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

```{r fig.height=4, fig.width=7}

simulation |>
  mutate(population_d = paste0("Population Cohen's d = ", pop_m2)) |>
  ggplot(aes(p_naive_p_diff)) +
  geom_histogram(binwidth = 0.0025, boundary = 0) +
  theme_linedraw() +
  ggtitle("Imprecision of p-values calculated from summary statistics\nwithout accounting for rounding") +
  facet_wrap(~ population_d)

```

## Example of false positive

```{r}

fp_examples <- simulation |>
  filter(!p_inbounds_reappraised) |>
  arrange(desc(p_abs_beyond_bounds_reappraised)) |>
  slice(1:5) |>
  unnest(results_original, names_sep = "_") |>
  select(
    m1 = results_original_m1,
    m2 = results_original_m2,
    sd1 = results_original_sd1,
    sd2 = results_original_sd2,
    n1 = results_original_n1,
    n2 = results_original_n2,
    p = results_original_p,
    reappraised_p_min = p_min_reappraised,
    reappraised_p_max = p_max_reappraised,
    reappraised_p_inbounds = p_inbounds_reappraised,
    analysis_rounding
  )

if (nrow(fp_examples) > 0) {
  fp_examples |>
    kable() |>
    kable_classic(full_width = FALSE)
} else {
  cat("No false positives found - reappraised achieved 100% coverage.")
}

```
