---
title: "Validation of {reappraised} for categorical (count) data"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r include=FALSE}

# formatting options
# set default chunk options
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

# disable scientific notation
options(scipen = 999)

```

# Dependencies

```{r}

library(tibble)
library(tidyr)
library(dplyr)
library(purrr)
library(furrr)
library(readr)
library(ggplot2)
library(scales)
library(knitr)
library(kableExtra)
library(reappraised)

```

# Functions

```{r}

generate_contingency_table <- function(p1, p2, n1, n2) {
  y1 <- rbinom(n1, 1, p1)
  y2 <- rbinom(n2, 1, p2)

  # 2x2 matrix: rows = outcome (0/1), cols = group (1/2)
  tab <- matrix(c(sum(y1 == 1), sum(y2 == 1),
                  sum(y1 == 0), sum(y2 == 0)),
                nrow = 2, byrow = TRUE)

  return(tab)
}

compute_reference_pvalues <- function(tab) {
  tibble(
    p_chi_ref       = suppressWarnings(chisq.test(tab, correct = FALSE)$p.value),
    p_chi_yates_ref = suppressWarnings(chisq.test(tab, correct = TRUE)$p.value),
    p_fisher_ref    = fisher.test(tab)$p.value
  )
}

compute_reappraised_pvalues <- function(tab) {
  p_fisher <- reappraised:::fisher_fast(tab)

  chi_no_corr <- reappraised:::chi_fast(tab, correct = FALSE)
  chi_yates   <- reappraised:::chi_fast(tab, correct = TRUE)

  tibble(
    p_chi           = chi_no_corr$p,
    p_chi_yates     = chi_yates$p,
    p_fisher        = p_fisher,
    p_lr            = reappraised:::lr(tab),
    p_cmh           = reappraised:::cmh(tab),
    p_midp_exact    = reappraised:::midp_exact(tab),
    p_midp_classic  = reappraised:::midp_classic(tab, p_fisher),
    chi_warning     = chi_no_corr$warning
  )
}

run_single_sim <- function(p1, p2, n1, n2) {
  tab <- generate_contingency_table(p1, p2, n1, n2)

  # Flag degenerate tables (zero row/column sums)
  if (any(rowSums(tab) == 0) || any(colSums(tab) == 0)) {
    return(tibble(
      a = tab[1,1], b = tab[1,2], c = tab[2,1], d = tab[2,2],
      degenerate = TRUE
    ))
  }

  ref   <- compute_reference_pvalues(tab)
  reapp <- compute_reappraised_pvalues(tab)

  bind_cols(
    tibble(a = tab[1,1], b = tab[1,2], c = tab[2,1], d = tab[2,2],
           degenerate = FALSE),
    ref,
    reapp
  )
}

```

# Run simulation

```{r}

if (file.exists("simulation_count.rds")) {
  simulation <- read_rds("simulation_count.rds")
} else {

  # simulation parameters
  experiment_parameters <- expand_grid(
    pop_p1 = c(0.1, 0.3, 0.5),
    pop_p2 = c(0.1, 0.3, 0.5, 0.7, 0.9),
    pop_n1 = c(10, 25, 50, 100),
    pop_n2 = c(10, 25, 50, 100),
    iteration = 1:500
  )

  # set up parallelization
  plan(multisession, workers = future::availableCores())

  # set seed
  set.seed(42)

  # run simulation
  simulation <- experiment_parameters |>
    mutate(results = future_pmap(
      .l = list(p1 = pop_p1, p2 = pop_p2, n1 = pop_n1, n2 = pop_n2),
      .f = run_single_sim,
      .progress = TRUE,
      .options = furrr_options(seed = TRUE)
    )) |>
    unnest(results)

  write_rds(simulation, "simulation_count.rds", compress = "gz")

}

```

# Degenerate tables

Tables with zero row or column margins (all successes or all failures in a group) cannot be tested. These are excluded from subsequent analyses.

```{r}

simulation |>
  group_by(pop_p1, pop_p2, pop_n1, pop_n2) |>
  summarize(
    n_total = n(),
    n_degenerate = sum(degenerate),
    pct_degenerate = mean(degenerate) * 100
  ) |>
  filter(n_degenerate > 0) |>
  kable(digits = 1) |>
  kable_classic(full_width = FALSE)

```

```{r}

# filter to non-degenerate for all subsequent analyses
simulation <- simulation |>
  filter(!degenerate) |>
  mutate(
    # absolute differences
    diff_chi       = abs(p_chi - p_chi_ref),
    diff_chi_yates = abs(p_chi_yates - p_chi_yates_ref),
    diff_fisher    = abs(p_fisher - p_fisher_ref),

    # exact matches (within floating point tolerance)
    match_chi       = near(p_chi, p_chi_ref, tol = 1e-10),
    match_chi_yates = near(p_chi_yates, p_chi_yates_ref, tol = 1e-10),
    match_fisher    = near(p_fisher, p_fisher_ref, tol = 1e-10),

    # expected cell warning flag
    has_chi_warning = (chi_warning != "")
  )

```

# Overall agreement rates

Do reappraised's internal functions produce identical p-values to R's built-in `chisq.test()` and `fisher.test()`?

```{r}

simulation |>
  summarize(
    n_sims               = n(),
    match_rate_chi        = mean(match_chi),
    match_rate_chi_yates  = mean(match_chi_yates),
    match_rate_fisher     = mean(match_fisher),
    max_diff_chi          = max(diff_chi),
    max_diff_chi_yates    = max(diff_chi_yates),
    max_diff_fisher       = max(diff_fisher)
  ) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

# Agreement by conditions

## By sample size

```{r}

simulation |>
  group_by(pop_n1, pop_n2) |>
  summarize(
    n = n(),
    match_rate_chi       = mean(match_chi),
    match_rate_chi_yates = mean(match_chi_yates),
    match_rate_fisher    = mean(match_fisher)
  ) |>
  kable(digits = 4) |>
  kable_classic(full_width = FALSE)

```

## By chi-squared expected cell warning

```{r}

simulation |>
  group_by(has_chi_warning) |>
  summarize(
    n = n(),
    match_rate_chi       = mean(match_chi),
    match_rate_chi_yates = mean(match_chi_yates),
    match_rate_fisher    = mean(match_fisher)
  ) |>
  kable(digits = 4) |>
  kable_classic(full_width = FALSE)

```

## By population proportions

```{r}

simulation |>
  group_by(pop_p1, pop_p2) |>
  summarize(
    n = n(),
    match_rate_chi       = mean(match_chi),
    match_rate_chi_yates = mean(match_chi_yates),
    match_rate_fisher    = mean(match_fisher)
  ) |>
  kable(digits = 4) |>
  kable_classic(full_width = FALSE)

```

# Scatter plots: reappraised vs. R built-in

```{r fig.height=5, fig.width=5}

simulation |>
  ggplot(aes(x = p_chi_ref, y = p_chi, color = match_chi)) +
  geom_abline(slope = 1, linetype = "dotted") +
  geom_point(alpha = 0.1, size = 0.5) +
  scale_color_viridis_d(begin = 0.2, end = 0.5) +
  labs(x = "R chisq.test (reference)", y = "reappraised chi_fast",
       title = "Chi-squared test (no correction)") +
  theme_linedraw()

simulation |>
  ggplot(aes(x = p_chi_yates_ref, y = p_chi_yates, color = match_chi_yates)) +
  geom_abline(slope = 1, linetype = "dotted") +
  geom_point(alpha = 0.1, size = 0.5) +
  scale_color_viridis_d(begin = 0.2, end = 0.5) +
  labs(x = "R chisq.test with Yates (reference)", y = "reappraised chi_fast with Yates",
       title = "Chi-squared test (Yates correction)") +
  theme_linedraw()

simulation |>
  ggplot(aes(x = p_fisher_ref, y = p_fisher, color = match_fisher)) +
  geom_abline(slope = 1, linetype = "dotted") +
  geom_point(alpha = 0.1, size = 0.5) +
  scale_color_viridis_d(begin = 0.2, end = 0.5) +
  labs(x = "R fisher.test (reference)", y = "reappraised fisher_fast",
       title = "Fisher's exact test") +
  theme_linedraw()

```

# Cross-test comparison

Distribution of p-values by test type across all simulations.

```{r fig.height=8, fig.width=7}

simulation |>
  slice_sample(n = min(5000, nrow(simulation))) |>
  select(p_chi, p_chi_yates, p_fisher, p_lr, p_cmh, p_midp_exact, p_midp_classic) |>
  pivot_longer(everything(), names_to = "test", values_to = "p_value") |>
  mutate(test = factor(test, levels = c("p_chi", "p_chi_yates", "p_fisher",
                                         "p_lr", "p_cmh", "p_midp_exact", "p_midp_classic"))) |>
  ggplot(aes(x = p_value)) +
  geom_histogram(binwidth = 0.02, boundary = 0) +
  facet_wrap(~ test, ncol = 2, scales = "free_y") +
  theme_linedraw() +
  labs(title = "Distribution of p-values by test type",
       x = "p-value", y = "Count")

```

Pairwise comparison: chi-squared vs Fisher's exact.

```{r fig.height=5, fig.width=5}

simulation |>
  slice_sample(n = min(5000, nrow(simulation))) |>
  ggplot(aes(x = p_chi, y = p_fisher)) +
  geom_abline(slope = 1, linetype = "dotted") +
  geom_point(alpha = 0.1, size = 0.5) +
  labs(x = "Chi-squared (no correction)", y = "Fisher's exact",
       title = "Chi-squared vs Fisher's exact") +
  theme_linedraw()

simulation |>
  slice_sample(n = min(5000, nrow(simulation))) |>
  ggplot(aes(x = p_chi, y = p_midp_exact)) +
  geom_abline(slope = 1, linetype = "dotted") +
  geom_point(alpha = 0.1, size = 0.5) +
  labs(x = "Chi-squared (no correction)", y = "Mid-p exact",
       title = "Chi-squared vs Mid-p exact") +
  theme_linedraw()

```

# Test disagreement

Cases where different tests lead to different significance conclusions at alpha = .05.

```{r}

test_disagreement <- simulation |>
  mutate(
    sig_chi       = p_chi < 0.05,
    sig_chi_yates = p_chi_yates < 0.05,
    sig_fisher    = p_fisher < 0.05,
    sig_midp      = p_midp_exact < 0.05,
    disagree_chi_fisher = sig_chi != sig_fisher,
    disagree_chi_yates  = sig_chi != sig_chi_yates,
    disagree_chi_midp   = sig_chi != sig_midp
  )

test_disagreement |>
  group_by(pop_p1, pop_p2) |>
  summarize(
    n = n(),
    pct_disagree_chi_fisher = mean(disagree_chi_fisher) * 100,
    pct_disagree_chi_yates  = mean(disagree_chi_yates) * 100,
    pct_disagree_chi_midp   = mean(disagree_chi_midp) * 100
  ) |>
  filter(pct_disagree_chi_fisher > 0 | pct_disagree_chi_yates > 0 | pct_disagree_chi_midp > 0) |>
  kable(digits = 2) |>
  kable_classic(full_width = FALSE)

```

By sample size:

```{r}

test_disagreement |>
  group_by(pop_n1, pop_n2) |>
  summarize(
    n = n(),
    pct_disagree_chi_fisher = mean(disagree_chi_fisher) * 100,
    pct_disagree_chi_yates  = mean(disagree_chi_yates) * 100,
    pct_disagree_chi_midp   = mean(disagree_chi_midp) * 100
  ) |>
  kable(digits = 2) |>
  kable_classic(full_width = FALSE)

```

# Expected cell count warnings

Frequency of the chi-squared expected-cell-count < 5 warning, by sample size and population proportion.

```{r}

simulation |>
  group_by(pop_p1, pop_p2, pop_n1) |>
  summarize(
    n = n(),
    pct_warning = mean(has_chi_warning) * 100
  ) |>
  filter(pct_warning > 0) |>
  kable(digits = 1) |>
  kable_classic(full_width = FALSE)

```
