---
title: "Getting started with recalc"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with recalc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 4
)
options(scipen = 999)
```

```{r setup}
library(recalc)
```

## The idea

A reported statistical result is not a single number. The inputs it was
computed from are rounded, and the analysis that produced it involved choices
that a paper usually does not state (Student's vs Welch's *t*, one- vs
two-sided, and so on). `recalc` recomputes the result across the *multiverse* of
those choices and reports the resulting interval, then checks whether the value
printed in the paper falls inside it.

## Independent-samples *t*-test

Given the reported group means, SDs, and *N*s (with the number of decimals each
was reported to), `recalc_independent_t()` returns the range of *p*-values and
Cohen's *d* values compatible with those inputs, and compares them against the
reported *p* and *d*.

```{r indep-t}
res <- recalc_independent_t(
  m1 = 10.3, m2 = 8.7,
  sd1 = 3.1, sd2 = 2.8,
  n1 = 50,   n2 = 48,
  m_digits  = 1,
  sd_digits = 1,
  d = 0.55, d_ci_lower = 0.20, d_ci_upper = 0.90, d_digits = 2,
  p = 0.0024, p_digits = 4
)

res$reproduced
```

The `reproduced` row carries the key verdicts: `p_inbounds` and `d_inbounds`
say whether the reported *p* and *d* fall inside the recalculated intervals,
together with the (rounded) interval bounds.

If you only care about one side, `recalc_independent_t_p()` and
`recalc_independent_t_d()` return the *p*-value and effect-size multiverses on
their own.

### Visualising the multiverse

`plot_multiverse_p()` and `plot_multiverse_d()` show every recalculated value
against the reported one, so a result that is only reproducible under a narrow
set of choices is easy to spot.

```{r plots}
plot_multiverse_p(res)
plot_multiverse_d(res)
```

### SD / SE confusion

A common source of irreproducible *t*-tests is a table that labels standard
errors as standard deviations. Setting `include_se_sd_confusion = TRUE` adds a
branch to the multiverse in which the reported SDs are reinterpreted as SEs and
converted with `SD = SE * sqrt(n)`.

```{r se-sd}
res_se <- recalc_independent_t(
  m1 = 10.3, m2 = 8.7,
  sd1 = 3.1, sd2 = 2.8,
  n1 = 50,   n2 = 48,
  m_digits  = 1,
  sd_digits = 1,
  p = 0.0024, p_digits = 4,
  include_se_sd_confusion = TRUE
)

res_se$reproduced[, c("p_inbounds", "min_p_rounded", "max_p_rounded")]
```

## Contingency tables

`recalc_chisq()` does the same for a chi-squared / association test. Give it the
counts (or reconstructed counts from reported proportions and group sizes) and,
optionally, the reported *p*; it recomputes the *p*-value across test variants
(Pearson, Yates, likelihood ratio, Fisher, and others) and checks the reported
value against the interval.

```{r chisq}
counts <- matrix(c(30, 20,
                   15, 35),
                 nrow = 2, byrow = TRUE)

res_chisq <- recalc_chisq(counts = counts, p = 0.002, p_digits = 3)

res_chisq$reproduced[, c("p_inbounds", "min_p_rounded", "max_p_rounded")]
```

## Other checks

The package includes further `recalc_*()` families, each following the same
report-an-interval-and-check pattern:

- `recalc_regression_from_cor_mat()` refits an OLS regression from a reported
  correlation matrix and recomputes coefficients, CIs, and *p*-values.
- `recalc_mediation()` recalculates indirect-effect estimates and *p*-values.
- `recalc_missing_subgroup()` checks whole-sample against subgroup descriptives.

See the function reference (`help(package = "recalc")`) for the full list.
