---
title: "Validation of {recalc} and {MEPtoolkit}"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

# TODO package

- add one vs two sided p value?
- other versions of d?
- input assumes both Ms and both SDs have identical precision. is this ever violated?
- t = results_original_t, # bounds?
- df = results_original_df) # value?
- consider using scrutiny::restore_zeros in min and max output, using m_digits, sd_digits, and d_digits

# TODO sims

- add welch's test and unbalanced SDs
- 

```{r include=FALSE}

# formatting options
# set default chunk options
knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE)

# disable scientific notation
options(scipen = 999) 

```

# Dependencies

```{r}

library(tibble)
library(roundwork)
# devtools::install_github("ianhussey/recalc")
library(recalc)
library(MEPtoolkit)

library(tidyr)
library(dplyr)
library(forcats)
library(readr)
library(purrr) 
library(furrr)
library(ggplot2)
library(ggstance)
library(scales)
library(knitr)
library(kableExtra)
library(parameters)
library(effectsize)

```

# Functions

```{r}

# functions for simulation
generate_data <- function(m1,
                          m2,
                          sd1,
                          sd2,
                          n1,
                          n2) {
  
  data1 <- 
    tibble(condition = "1",
           score = rnorm(n = n1, mean = m1, sd = sd1))
  
  data2 <- 
    tibble(condition = "2",
           score = rnorm(n = n2, mean = m2, sd = sd2))
  
  data_combined <- bind_rows(data1,
                             data2)
  
  return(data_combined)
}

analyze <- function(data, 
                    m_digits,
                    sd_digits,
                    d_digits,
                    p_digits,
                    analysis_rounding) {
  
  res_summary <- data |>
    group_by(condition) |>
    summarize(m = mean(score),
              sd = sd(score),
              n = n()) |>
    pivot_wider(names_from = condition, 
                names_sep = "",
                values_from = c("m", "sd", "n"))
  
  res_ttest <- 
    t.test(formula = score ~ condition, 
           data = data,
           var.equal = TRUE,
           alternative = "two.sided") |>
    model_parameters() |>
    as_tibble() |>
    select(t, df = df_error, p)
  
  res_d <- effectsize::cohens_d(score ~ condition,
                                data = data) |>
    as_tibble() |>
    select(d = Cohens_d,
           d_ci_lower = CI_low,
           d_ci_upper = CI_high)
  
  res <- bind_cols(res_summary, 
                   res_ttest,
                   res_d)
  
  if(analysis_rounding == "half_up"){
    res <- res |>
      mutate(m1 = round_up(m1, m_digits),
             m2 = round_up(m2, m_digits),
             sd1 = round_up(sd1, sd_digits),
             sd2 = round_up(sd2, sd_digits),
             d = round_up(d, d_digits),
             d_ci_lower = round_up(d_ci_lower, d_digits),
             d_ci_upper = round_up(d_ci_upper, d_digits),
             p = round_up(p, p_digits),
             p_operator = dplyr::if_else(dplyr::near(p, 0), "less_than", "equals"),
             p = dplyr::if_else(dplyr::near(p, 0), 1 * 10^-p_digits, p))
  } else if(analysis_rounding == "half_down"){
    res <- res |>
      mutate(m1 = round_down(m1, m_digits),
             m2 = round_down(m2, m_digits),
             sd1 = round_down(sd1, sd_digits),
             sd2 = round_down(sd2, sd_digits),
             d = round_down(d, d_digits),
             d_ci_lower = round_down(d_ci_lower, d_digits),
             d_ci_upper = round_down(d_ci_upper, d_digits),
             p = round_down(p, p_digits),
             p_operator = dplyr::if_else(dplyr::near(p, 0), "less_than", "equals"),
             p = dplyr::if_else(dplyr::near(p, 0), 1 * 10^-p_digits, p))
  } else if(analysis_rounding == "bankers"){
    res <- res |>
      mutate(m1 = round(m1, m_digits),
             m2 = round(m2, m_digits),
             sd1 = round(sd1, sd_digits),
             sd2 = round(sd2, sd_digits),
             d = round(d, d_digits),
             d_ci_lower = round(d_ci_lower, d_digits),
             d_ci_upper = round(d_ci_upper, d_digits),
             p = round(p, p_digits),
             p_operator = dplyr::if_else(dplyr::near(p, 0), "less_than", "equals"),
             p = dplyr::if_else(dplyr::near(p, 0), 1 * 10^-p_digits, p))
  } else if(analysis_rounding == "trunc"){
    res <- res |>
      mutate(m1 = round_trunc(m1, m_digits),
             m2 = round_trunc(m2, m_digits),
             sd1 = round_trunc(sd1, sd_digits),
             sd2 = round_trunc(sd2, sd_digits),
             d = round_trunc(d, d_digits),
             d_ci_lower = round_trunc(d_ci_lower, d_digits),
             d_ci_upper = round_trunc(d_ci_upper, d_digits),
             p = round_trunc(p, p_digits),
             p_operator = dplyr::if_else(dplyr::near(p, 0), "less_than", "equals"),
             p = dplyr::if_else(dplyr::near(p, 0), 1 * 10^-p_digits, p))
  } 
  
  return(res)
}

reconstruct <- function(results, 
                        m_digits,
                        sd_digits,
                        d_digits,
                        p_digits,
                        direction,
                        hedges_correction,
                        include_se_sd_confusion,
                        reconstruct_rounding,
                        output_rounding){
  independent_t_test_summary(
    m1 = results$m1,
    m2 = results$m2,
    sd1 = results$sd1,
    sd2 = results$sd2,
    n1 = results$n1,
    n2 = results$n2,
    d = results$d,
    d_ci_lower = results$d_ci_lower,
    d_ci_upper = results$d_ci_upper,
    p_operator = results$p_operator,
    p = results$p,
    input_rounding = reconstruct_rounding,
    output_rounding = output_rounding,
    m_digits = m_digits,
    sd_digits = sd_digits,
    d_digits = d_digits,
    direction = direction,
    hedges_correction = hedges_correction,
    p_digits = p_digits,
    include_se_sd_confusion = include_se_sd_confusion
  )$reproduced
}



#' @export
reconstruct_t_slicer <- function(results, output = FALSE){
  
  fit <- t_slicer(m1 = as.character(results$m1),
                  s1 = as.character(results$sd1),
                  n1 = results$n1,
                  m2 = as.character(results$m2),
                  s2 = as.character(results$sd2),
                  n2 = results$n2,
                  p  = as.character(results$p),
                  output = output)
  
  res <- data.frame(
    tslicer_p_rivits_student = fit$Student_RIVETS,
    tslicer_p_min_student = fit$Student_min,
    tslicer_p_max_student = fit$Student_max,
    tslicer_p_min_welch = fit$Welch_min,
    tslicer_p_max_welch = fit$Welch_max
  ) |>
    mutate_all(as.numeric) |>
    mutate(p_inbounds_tslicer = if_else(between(results$p, tslicer_p_min_student, tslicer_p_max_student) |
                                  between(results$p, tslicer_p_min_welch, tslicer_p_max_welch), 
                                TRUE,
                                FALSE))
  
  return(res)
}

```

# Run simulation

```{r}

if(file.exists("simulation.rds")){
  simulation <- read_rds("simulation_mep.rds")
} else {
  
  # simulation parameters
  experiment_parameters <- expand_grid(
    pop_m1 = 0,
    pop_m2 = c(0, 2),
    pop_sd1 = 1,
    pop_sd2 = 1, 
    pop_n1 = 25,
    pop_n2 = 25, 
    m_digits = 2,
    sd_digits = 2,
    d_digits = 2,
    p_digits = 3,
    analysis_rounding = c("half_up", "half_down", "bankers", "trunc"),
    reconstruct_rounding = c("truncated", "rounded"),
    output_rounding = c("half_up", "half_down", "bankers", "trunc"),
    include_se_sd_confusion = FALSE,
    direction = "m1_minus_m2",
    hedges_correction = FALSE, 
    iteration = 1:100
  ) 
  
  # set up parallelization
  plan(multisession, workers = future::availableCores())
  
  # set seed
  set.seed(42)
  
  # run simulation
  simulation <- experiment_parameters |>
    # generate datasets
    mutate(generated_data = future_pmap(.l = list(m1 = pop_m1,
                                                  m2 = pop_m2,
                                                  sd1 = pop_sd1,
                                                  sd2 = pop_sd2,
                                                  n1 = pop_n1,
                                                  n2 = pop_n2),
                                        .f = generate_data,
                                        .progress = TRUE,
                                        .options = furrr_options(seed = TRUE))) |>
    # analyze data to create the real p and d values and summary stats
    mutate(results_original = future_pmap(.l = list(data = generated_data,
                                                    m_digits = m_digits,
                                                    sd_digits = sd_digits,
                                                    d_digits = d_digits,
                                                    p_digits = p_digits,
                                                    analysis_rounding = analysis_rounding),
                                          .f = analyze,
                                          .progress = TRUE,
                                          .options = furrr_options(seed = TRUE))) |>
    # use recalc to reconstruct the p and d bounds
    mutate(results_reconstructed = future_pmap(.l = list(results = results_original,
                                                         m_digits = m_digits,
                                                         sd_digits = sd_digits,
                                                         d_digits = d_digits,
                                                         p_digits = p_digits,
                                                         direction = direction,
                                                         hedges_correction = hedges_correction,
                                                         include_se_sd_confusion = include_se_sd_confusion,
                                                         reconstruct_rounding = reconstruct_rounding,
                                                         output_rounding = output_rounding),
                                               .f = reconstruct,
                                               .progress = TRUE,
                                               .options = furrr_options(seed = TRUE))) |>
    unnest(results_reconstructed) |>
    mutate(d_beyond_bounds = case_when(d < min_d_rounded ~ d-min_d_rounded,
                                       d > max_d_rounded ~ max_d_rounded-d,
                                       TRUE ~ 0),
           d_abs_beyond_bounds = abs(d_beyond_bounds),
           p_beyond_bounds = case_when(p < min_p_rounded ~ p-min_p_rounded,
                                       p > max_p_rounded ~ max_p_rounded-p,
                                       TRUE ~ 0),
           p_abs_beyond_bounds = abs(p_beyond_bounds)) |>
    
    # use reappraised to reconstruct the p bounds
    mutate(results_reconstructed_t_slicer = future_pmap(.l = list(results = results_original),
                                                       .f = reconstruct_t_slicer,
                                                       .progress = TRUE,
                                                       .options = furrr_options(seed = TRUE))) |>
    unnest(results_reconstructed_t_slicer) 
  
  write_rds(simulation, "simulation_mep.rds", compress = "gz")
  
}

```

# Summarize results

## recalc's coverage

All results

```{r}

simulation_coverage <- simulation |>
  filter(pop_m2 == 0) |> # hack for the moment
  #group_by(pop_es = pop_m2, analysis_rounding, reconstruct_rounding, output_rounding) |>
  summarize(
    coverage_p_recalc = mean(p_inbounds),
    coverage_p_tslicer = mean(p_inbounds_tslicer)
  )

simulation_coverage |>
  kable() |>
  kable_classic(full_width = FALSE)

```

By factorial, only cells with imperfect coverage

```{r}

simulation_coverage_by <- simulation |>
  filter(pop_m2 == 0) |> # hack for the moment
  group_by(pop_es = pop_m2, analysis_rounding, reconstruct_rounding, output_rounding) |>
  summarize(
    coverage_p_recalc = mean(p_inbounds),
    coverage_p_tslicer = mean(p_inbounds_tslicer)
  )

simulation_coverage_by |>
  filter(!near(coverage_p_recalc, 1) |
           !near(coverage_p_tslicer, 1)) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

By factorial, only cells with imperfect coverage, output_rounding != "trunc"

```{r}

simulation_coverage_by |>
  filter(output_rounding != "trunc") |>
  filter(!near(coverage_p_recalc, 1) |
           !near(coverage_p_tslicer, 1)) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

By factorial, only cells with imperfect coverage, output_rounding != "trunc" & reconstruct_rounding == "truncated"


Imperfect coverage is expected when the analysis was reported using truncation (offset up to ±.01) but the reconstruction was done using rounding (offset up to ±.005):

## MEPtoolkit t_slicer's coverage

only output_rounding == "half_up"

```{r}

simulation |>
  filter(pop_m2 == 0) |> # hack for the moment
  filter(output_rounding == "half_up") |>
  rowwise() |>
  mutate(p_min_tslicer = min(c(tslicer_p_min_student, tslicer_p_min_welch)),
         p_max_tslicer = min(c(tslicer_p_max_student, tslicer_p_max_welch))) |>
  ungroup() |>
  #filter(!p_inbounds_tslicer) |>
  ggplot(aes(x = p, y = tslicer_p_rivits_student, color = p_inbounds_tslicer)) +
  geom_abline(slope = 1, linetype = "dotted") +
  geom_linerange(aes(ymin = p_min_tslicer, ymax = p_max_tslicer), alpha = 0.6) +
  geom_point(alpha = 0.6) +
  theme_linedraw() +
  facet_grid(analysis_rounding ~ reconstruct_rounding) +
  ggtitle("MEPtoolkit's false positives")

simulation |>
  filter(pop_m2 == 0) |> # hack for the moment
  filter(output_rounding == "half_up") |>
  rowwise() |>
  mutate(p_min_tslicer = min(c(tslicer_p_min_student, tslicer_p_min_welch)),
         p_max_tslicer = min(c(tslicer_p_max_student, tslicer_p_max_welch))) |>
  ungroup() |>
  filter(!p_inbounds_tslicer) |>
  ggplot(aes(x = p, y = tslicer_p_rivits_student, color = p_inbounds_tslicer)) +
  geom_abline(slope = 1, linetype = "dotted") +
  geom_linerange(aes(ymin = p_min_tslicer, ymax = p_max_tslicer), alpha = 0.6) +
  geom_point(alpha = 0.6) +
  theme_linedraw() +
  facet_grid(analysis_rounding ~ reconstruct_rounding) +
  ggtitle("MEPtoolkit's false positives")

```

### Example of a dataset that generates an incorrect conclusion (false positive inconsistency) 

TODO

```{r}

# reappraised_example <- simulation |>
#   filter(analysis_rounding == "half_up" & reconstruct_rounding == "rounded") |>
#   arrange(desc(p_abs_beyond_bounds_bolland)) |>
#   slice(1:1) |>
#   unnest(results_original, names_sep = "_") |>
#   select(m1 = results_original_m1,
#          m2 = results_original_m2,
#          sd1 = results_original_sd1,
#          sd2 = results_original_sd2,
#          n1 = results_original_n1,
#          n2 = results_original_n2,
#          p = results_original_p,
#          recalc_p_min_rounded = min_p_rounded,
#          recalc_p_max_rounded = max_p_rounded,
#          recalc_p_inbounds = p_inbounds,
#          bolland_p_min = p_min_bolland,
#          bolland_p_max = p_max_bolland,
#          bolland_p_inbounds = p_inbounds_tslicer,
#          analysis_rounding)
# 
# reappraised_example |>
#   kable() |>
#   kable_classic(full_width = FALSE)
# 
# write_csv(reappraised_example, "reappraised_example.csv")
# 
# 
# 
# reappraised_example_data <- simulation |>
#   filter(analysis_rounding == "half_up" & reconstruct_rounding == "rounded") |>
#   arrange(desc(p_abs_beyond_bounds_bolland)) |>
#   slice(1:1) |>
#   select(generated_data) |>
#   unnest(generated_data) 
# 
# reappraised_example_data |>
#   kable() |>
#   kable_classic(full_width = FALSE)
# 
# write_csv(reappraised_example_data, "reappraised_example_data.csv")
# 
# reappraised_example_data |>
#   group_by(condition) |>
#   summarize(n = n(),
#             mean = mean(score),
#             sd = sd(score)) |>
#   mutate_if(is.numeric, roundwork::round_down, digits = 2) |>
#   kable() |>
#   kable_classic(full_width = FALSE)

```




