---
title: "Head-to-head comparison of `recalc` and `reappraised` for categorical p-values via simulation"
author: "Ian Hussey"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc_float: true
    code-fold: true
    code-overflow: wrap
    code-tools: true
    warning: false
    df-print: kable
    theme:
      light: cosmo
      dark: darkly
    highlight-style:
      light: breeze
      dark: dracula
    include-in-header:
      text: |
        <style>
          .kable-table table { width: 70%; margin: 0 auto; }
          .quarto-color-scheme-toggle::after { content: " Dark mode"; font-family: sans-serif; font-size: 0.85em; }
        </style>
---

# TODO

- latest app is ahead of the package, and doesn't rely on the package. might need a further comparison, with code extracted from the app.

```{r, include=FALSE}
# settings: hidden from rendered file
# disable messages and warnings in rendered files
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

# disable scientific numbers in output
options(scipen = 999)
```

# Introduction

<!-- Specify the simulation that will be run here. -->

This introduction is structured following the ADEMP framework (Aims, Data-generating process, Estimands, Methods, Performance) as described by Morris, White, and Crowther (2019) and operationalised for psychology by Siepe et al. (2024).

## Project description

<!-- Provide a brief description of the project, including any empirical examples. -->

This file runs `recalc::recalc_chisq_p()` and `reappraised`'s categorical machinery on identical simulated 2×2 tables and quantifies the **coverage of each tool when the analyst's test method is unknown**.

- **`recalc::recalc_chisq_p()`** — current version of `recalc`.
- **`reappraised`** — the per-table categorical *p*-value computation from `R/baseline_pvalues_cat.R` of the package (`pval_cat_fn()`, Bolland et al.), inlined verbatim below.

**`reappraised` version pinned for this simulation:**

- Package: `reappraised` **v0.2**
- Source: **CRAN** (`Repository: CRAN`, `RemoteSha: 0.2`)
- Package date: **2026-03-05**; built under R 4.5.2.

The package version actually loaded is printed at run time in the *Reproducibility* section below and in `sessionInfo()`, so the rendered document always records the version it was built against.

### Why coverage across analytic choices, not point reproduction

The same 2×2 table yields materially different *p*-values depending on which test a statistical package computed: Pearson chi-square (no correction), Yates' continuity-corrected chi-square, the likelihood-ratio G², Fisher's exact test, or one of several mid-*p* variants. **In practice the analyst reading a published table usually does not know which of these produced the reported *p*.** A consistency-checking tool must therefore behave as a *multiverse / interval estimator*: it hedges across the methods it knows and asks whether the reported value falls inside the resulting `[min_p, max_p]`.

The two tools span different method sets:

- `reappraised`'s `pval_cat_fn()` implements **four** tests, selected by its `stat` argument: `"chisq"` (Pearson, `correct = FALSE`), `"fisher"` (Fisher's exact), `"midp"` / `"midp.epitools"` (central mid-*p* via `epitools::riskratio.wald()`), and `"midp.sas"` (SAS `PROC FREQ` exact mid-*p*). It has **no** Yates or likelihood-ratio option, and — because cell counts are exact integers — **no** rounding multiverse.
- `recalc::recalc_chisq_p()` spans **nine** methods (`pearson`, `yates`, `likelihood_ratio`, `n1_chisq`, `fisher`, `fisher_central`, `fisher_midp`, `fisher_midp_sas`, `wald_z`) and additionally offers an input-rounding route (reconstructing integer tables from rounded proportions).

So this simulation does **not** privilege either tool by treating one as ground truth. Instead, for each simulated table we generate a "reported" *p* under one of the **nine** analytic choices a real analyst could have made — using an **independent oracle** for each method (base R / closed-form / a separate mid-*p* implementation, never `recalc` or `reappraised` itself) — and then evaluate, for *both* tools, whether that reported *p* falls inside the multiverse the tool would build.

Three coverage series are reported throughout:

1. **`recalc`** — multiverse = span over its nine methods.
2. **`reappraised (span)`** — multiverse = span over the four tests it implements (the most favourable reading: `reappraised` used as a hedge across all of its options).
3. **`reappraised (default)`** — its single default `stat = "chi_midp"` (Pearson, falling back to mid-*p* when an expected cell is < 5), i.e. `reappraised` as typically used out of the box, producing one number.

The **failure modes** of interest are the five analytic choices outside `reappraised`'s four-test set: **Yates' correction** (R's `chisq.test()` default; SPSS "Continuity Correction"), the **likelihood ratio** (SPSS/SAS "Likelihood Ratio"), the **N−1 / Campbell** chi-square (SPSS "Linear-by-Linear"; SAS single-stratum Mantel-Haenszel), the **central-convention Fisher** exact, and the **unpooled Wald** two-proportion *z* — none of which `reappraised`'s categorical machinery can reproduce.

For the four shared tests, `recalc` and `reappraised` should agree numerically to floating point (verified in the fidelity section); the coverage differences therefore come entirely from method *breadth*, not from disagreement on shared methods.

**Which `reappraised` is this?** This head-to-head benchmarks the `reappraised` **package** function `pval_cat_fn()` (the four tests above; the "span" and single-`chi_midp` "default" readings). The `reappraised` **Shiny app** (`apps/app_dichotomous.R`) is a separate, more capable artifact and behaves *differently*: it computes a **battery of seven tests** — Pearson, Yates, likelihood ratio, single-stratum CMH/Mantel-Haenszel (verified identical to `recalc`'s `n1_chisq` = (N−1)/N × Pearson), Fisher (minlike), epitools central mid-*p*, and SAS classic mid-*p* — and its headline `any_match` flag reports whether the reported *p* **matches any** of them. That is a gap-aware **union**, like `recalc`'s default — *not* a single default and *not* a `[min, max]` span. Relative to `recalc`'s nine methods the app covers seven (it omits only `fisher_central` and `wald_z`), so on this axis the app is close to `recalc` and the "span"/"default" failure-mode results below characterise the **package function, not the app**.

## Prior related work

The companion file `validation/chi squared/simulation_p_recalc.qmd` validates `recalc_chisq_p()` against independent oracles (including `reappraised`'s own SAS mid-*p* algorithm). This file is the head-to-head against `reappraised` over a shared DGP, parallel to `validation/t-test/simulation_p_recalc_vs_reappraised.qmd` for the continuous case.

## Aims

1. **Numerical fidelity** — confirm `recalc` reproduces `reappraised`'s categorical *p*-value to floating-point tolerance for each of the four shared tests.
2. **Coverage across unknown analytic choices** — quantify, for `recalc` and for both readings of `reappraised`, the fraction of reports they cover when the report could have been produced by any of the nine tests.
3. **Quantify the failure modes** — measure how often `reappraised` misses reports produced by the five tests outside its set (Yates, likelihood ratio, N−1, central Fisher, Wald *z*), and how that varies with table sparsity.
4. **Coverage / width trade-off** — report each tool's *covered-region* width (for `recalc`, its default **union** width, not the convex hull) alongside its coverage, to show that broad coverage need not cost a wide consistent region.
5. **Rejection power / specificity** (second simulation) — sweep candidate reported *p*-values across the axis and quantify the proportion each rule correctly **rejects**, exposing the higher power of `recalc`'s gap-aware union over the hull-based rules.

## Data-Generating Process

### DGP specification approach

The DGP is **fully parametric**: for each replication a 2×2 table is formed from two independent Binomial draws (rows = groups, columns = success/failure),

$$X_j \overset{\text{ind}}{\sim} \text{Binomial}(n_j,\, \pi_j), \qquad j \in \{1, 2\}.$$

Small group sizes and probabilities away from the extremes produce sparse tables with low expected cell counts — the regime where the nine tests diverge most, and therefore where method breadth matters most.

### DGP factors

1. `n_per_group` — common per-group sample size.
2. `pi_1` — success probability in group 1.
3. `delta` — added to `pi_1` for group 2 (`pi_2 = pi_1 + delta`); `delta = 0` is the true null.

### Factor values and settings

- `n_per_group` ∈ {20, 80, 200}.
- `pi_1` ∈ {0.3, 0.5}.
- `delta` ∈ {0, 0.2, 0.4} (true null, moderate, and strong association). The `0.4` level is included to populate the small-*p* region so interval width — which depends on where on the *p*-axis the multiverse sits — can be stratified by *p*.

**Held constant:** balanced design.

### Analytic-choice factor

`analysis_method` ∈ {`"pearson"`, `"yates"`, `"likelihood_ratio"`, `"n1_chisq"`, `"fisher"`, `"fisher_central"`, `"fisher_midp"`, `"fisher_midp_sas"`, `"wald_z"`} — which test the analyst used to produce the reported *p*. The five methods outside `reappraised`'s set (`yates`, `likelihood_ratio`, `n1_chisq`, `fisher_central`, `wald_z`) are `reappraised`'s failure modes.

### Reporting-precision factor

`p_digits` ∈ {2, 3} — decimal places of the reported *p*-value, **swept** rather than fixed. (Cell counts are exact integers, so there is no proportion/percentage precision to sweep on the categorical side.)

### Factor combination and number of conditions

Fully factorial: 3 × 2 × 3 × 9 (DGP + analytic choice) × 2 (precision sweep) = **324 unique conditions**, each replicated K times.

## Estimands and Targets

Per replication: (a) for the fidelity check, `|p_recalc − p_reappraised|` on each shared test; (b) for coverage, the truth value of "reported *p* ∈ tool's `[min_p, max_p]`" for each of the three coverage series.

## Methods and extracted quantities

For each replication we (1) generate a 2×2 table; (2) compute the reported *p* under `analysis_method` with an **independent oracle**; (3) round it to `p_digits` (reporting `p < 10^(-p_digits)` when it collapses to 0); (4) evaluate inbounds for `recalc` (nine-method span), `reappraised (span)` (four-method span), and `reappraised (default)` (single `chi_midp` value), using a common rounded-bounds comparison.

## Performance and Uncertainty

### Performance measures

For each of the three coverage series (`recalc`, `reappraised (span)`, `reappraised (default)`) the primary measure is **empirical coverage**, the proportion of replications in a cell whose reported *p* falls inside that tool's interval:

$$\widehat{C} = \frac{1}{K} \sum_{k=1}^{K} \mathbb{1}\{p_k \in [\text{min}_k, \text{max}_k]\}.$$

Alongside coverage we report **interval width** (mean and max per cell, with MCSE for the mean), the **per-replication disagreement pattern** across the three series, and **fidelity** (max absolute cross-tool difference on the four shared tests).

### Monte Carlo uncertainty

Binomial MCSE for each coverage proportion: $\widehat{\text{MCSE}}(\widehat{C}) = \sqrt{\widehat{C}(1 - \widehat{C})/K}$. For the mean interval width, $\widehat{\text{MCSE}} = \text{sd(width)}/\sqrt{K}$.

### Number of simulation repetitions

K = 500 per condition for the full run (the file ships with a smaller K for fast iteration; see the experiment grid). This resolves the failure-mode coverage gaps (Yates, likelihood ratio) to well within 1%.

### Non-convergence and missing values

The only failure mode is a **degenerate table** (an empty margin, hence an undefined test). These are rare, yield `NA`, and are excluded case-wise. Both tools' per-table computations are otherwise closed-form/exact.

### Interpretation of performance measures

`recalc` should cover all nine analytic choices (every candidate test is in its span). `reappraised (span)` should cover its four implemented tests and miss a fraction of the five reports it cannot reproduce (`yates`, `likelihood_ratio`, `n1_chisq`, `fisher_central`, `wald_z`); `reappraised (default)` should cover only the analytic choices whose *p* rounds to its single `chi_midp` value. On the four shared tests the two tools should agree to floating-point tolerance.

# Methods

## Dependencies

```{r}
library(tidyr)
library(dplyr)
library(purrr)
library(furrr)
library(tibble)
library(readr)
library(ggplot2)
library(scales)
library(knitr)
library(kableExtra)
library(roundwork)
library(epitools)     # used by reappraised's mid-p calculation

# devtools::install_github("ianhussey/recalc")
library(recalc)

# Parallelisation — L'Ecuyer-CMRG stream via furrr_options(seed = TRUE).
plan(multisession, workers = future::availableCores())
```

### Reproducibility: pinned `reappraised` version

```{r}
reappraised_version <- as.character(utils::packageVersion("reappraised"))
reappraised_desc    <- utils::packageDescription("reappraised")
tibble::tibble(
  package    = "reappraised",
  version    = reappraised_version,
  date       = reappraised_desc$Date,
  repository = ifelse(is.null(reappraised_desc$Repository), NA, reappraised_desc$Repository),
  remote_sha = ifelse(is.null(reappraised_desc$RemoteSha),  NA, reappraised_desc$RemoteSha)
) |>
  kable(caption = "reappraised version used for this comparison") |>
  kable_styling(full_width = FALSE)

stopifnot(reappraised_version == "0.2")  # the version this file was written against
```

### `reappraised` internals (inlined)

The function below reproduces, **verbatim**, the per-table categorical *p*-value computation from
<https://github.com/cran/reappraised/blob/main/R/baseline_pvalues_cat.R>
(`pval_cat_fn()`, MIT-licensed, Bolland et al., v0.2). In the package this code runs inside a per-row loop over a wide summary data frame (the `method = "sm"` branch); here it is lifted into a single function taking a 2×2 matrix `x` and a `stat` string, with cell indexing and matrix orientation matching the original.

```{r}
# Per-table categorical p, lifted verbatim from reappraised::pval_cat_fn()
# (R/baseline_pvalues_cat.R, summary-data 'sm' branch). `x` is a 2x2 integer
# matrix with rows = groups, columns = (level 1, level 2).
reappraised_cat_p <- function(x, stat) {
  p_out <- NA_real_

  if (stat %in% c("chisq", "chi_fish", "chi_midp", "chi_midp.epi", "chi_midp.sas")) {
    chi   <- suppressWarnings(chisq.test(x, correct = FALSE))
    p_out <- chi$p.value
  }

  if (stat == "fisher") {
    if (sum(rowSums(x) == 0) > 0) {
      p_out <- 1
    } else {
      p_out <- suppressWarnings(fisher.test(x, workspace = 2e8)$p.value)
    }
  }

  if (stat %in% c("midp", "midp.epitools")) {
    p_out <- suppressWarnings(epitools::riskratio.wald(x)$p.value[2][1])
  }

  if (stat == "midp.sas") {
    d1 <- x[[1]]; d2 <- x[[2]]; d3 <- x[[3]]; d4 <- x[[4]]
    d5 <- d1 + d2 + d3 + d4
    num   <- c(seq(max(d1 + d2, 1), 1, -1), seq(max(d3 + d4, 1), 1, -1),
               seq(max(d1 + d3, 1), 1, -1), seq(max(d2 + d4, 1), 1, -1))
    denom <- c(seq(max(d5, 1), 1, -1), seq(max(d1, 1), 1, -1), seq(max(d2, 1), 1, -1),
               seq(max(d3, 1), 1, -1), seq(max(d4, 1), 1, -1))
    len <- length(num) - length(denom)
    if (len > 0) denom <- c(denom, rep(1, len))
    if (len < 0) num   <- c(num,   rep(1, abs(len)))
    p <- 100000
    num   <- sort(num,   decreasing = TRUE)
    denom <- sort(denom, decreasing = TRUE)
    for (j in seq_along(denom)) p <- p * num[j] / denom[j]
    p <- p / 100000
    p_out <- round(suppressWarnings(fisher.test(x, workspace = 2e8)$p.value) - 0.5 * p, 9)
  }

  as.numeric(p_out)
}

# The four tests reappraised implements.
reappraised_stats <- c("chisq", "fisher", "midp.epitools", "midp.sas")

# reappraised's multiverse: span over the four tests it can compute.
reappraised_span <- function(tab) {
  ps <- vapply(reappraised_stats,
               function(s) tryCatch(reappraised_cat_p(tab, s), error = function(e) NA_real_),
               numeric(1))
  ps <- ps[is.finite(ps)]
  if (length(ps) == 0) return(c(NA_real_, NA_real_))
  c(min(ps), max(ps))
}

# reappraised's DEFAULT single output: stat = "chi_midp" = Pearson, falling
# back to mid-p when an expected cell < 5 (the 2x2 rule from pval_cat_fn()).
reappraised_default <- function(tab) {
  E   <- outer(rowSums(tab), colSums(tab)) / sum(tab)
  chi <- suppressWarnings(chisq.test(tab, correct = FALSE))
  exp_low <- (length(E) == 4 && sum(E < 5) > 0)
  if (exp_low) reappraised_cat_p(tab, "midp.epitools") else chi$p.value
}
```

### Independent oracles (reported p per analytic choice, computed without either tool)

Each uses a different implementation path from both `recalc` and `reappraised`, so the reported *p* privileges neither tool.

```{r}
oracle_lr <- function(tab) {
  N <- sum(tab)
  if (any(rowSums(tab) == 0) || any(colSums(tab) == 0)) return(NA_real_)
  E  <- outer(rowSums(tab), colSums(tab)) / N
  nz <- tab > 0
  G  <- 2 * sum(tab[nz] * log(tab[nz] / E[nz]))
  pchisq(G, (nrow(tab) - 1L) * (ncol(tab) - 1L), lower.tail = FALSE)
}

oracle_midp_central <- function(tab) {
  r1 <- sum(tab[1, ]); r2 <- sum(tab[2, ]); c1 <- sum(tab[, 1]); N <- r1 + r2
  if (r1 == 0 || r2 == 0 || c1 == 0 || c1 == N) return(NA_real_)
  a       <- tab[1, 1]
  support <- max(0, c1 - r2):min(r1, c1)
  probs   <- dhyper(support, m = r1, n = r2, k = c1)
  p_obs   <- dhyper(a,       m = r1, n = r2, k = c1)
  min(1, 2 * min(sum(probs[support <= a]) - 0.5 * p_obs,
                 sum(probs[support >= a]) - 0.5 * p_obs))
}

oracle_midp_sas <- function(tab) {
  pf <- tryCatch(suppressWarnings(fisher.test(tab)$p.value), error = function(e) NA_real_)
  if (is.na(pf)) return(NA_real_)
  d1 <- tab[1, 1]; d2 <- tab[1, 2]; d3 <- tab[2, 1]; d4 <- tab[2, 2]; d5 <- d1 + d2 + d3 + d4
  num   <- c(seq(max(d1 + d2, 1), 1, -1), seq(max(d3 + d4, 1), 1, -1),
             seq(max(d1 + d3, 1), 1, -1), seq(max(d2 + d4, 1), 1, -1))
  denom <- c(seq(max(d5, 1), 1, -1), seq(max(d1, 1), 1, -1), seq(max(d2, 1), 1, -1),
             seq(max(d3, 1), 1, -1), seq(max(d4, 1), 1, -1))
  len <- length(num) - length(denom)
  if (len > 0) denom <- c(denom, rep(1, len))
  if (len < 0) num   <- c(num,   rep(1, abs(len)))
  num <- sort(num, decreasing = TRUE); denom <- sort(denom, decreasing = TRUE)
  p <- 100000; for (j in seq_along(denom)) p <- p * num[j] / denom[j]; p <- p / 100000
  round(pf - 0.5 * p, 9)
}

oracle_n1 <- function(tab) {
  if (any(rowSums(tab) == 0) || any(colSums(tab) == 0)) return(NA_real_)
  N  <- sum(tab)
  x2 <- suppressWarnings(chisq.test(tab, correct = FALSE)$statistic)
  pchisq((N - 1) / N * x2, (nrow(tab) - 1L) * (ncol(tab) - 1L), lower.tail = FALSE)
}
oracle_fisher_central <- function(tab) {
  r1 <- sum(tab[1, ]); r2 <- sum(tab[2, ]); c1 <- sum(tab[, 1]); N <- r1 + r2
  if (r1 == 0 || r2 == 0 || c1 == 0 || c1 == N) return(NA_real_)
  a <- tab[1, 1]; support <- max(0, c1 - r2):min(r1, c1)
  probs <- dhyper(support, m = r1, n = r2, k = c1)
  min(1, 2 * min(sum(probs[support <= a]), sum(probs[support >= a])))
}
oracle_wald_z <- function(tab) {
  n1 <- sum(tab[1, ]); n2 <- sum(tab[2, ])
  if (n1 == 0 || n2 == 0) return(NA_real_)
  p1 <- tab[1, 1] / n1; p2 <- tab[2, 1] / n2
  se <- sqrt(p1 * (1 - p1) / n1 + p2 * (1 - p2) / n2)
  if (!is.finite(se) || se == 0) return(NA_real_)
  2 * pnorm(-abs((p1 - p2) / se))
}

oracle_p <- function(tab, method) {
  out <- tryCatch(suppressWarnings(switch(
    method,
    pearson          = chisq.test(tab, correct = FALSE)$p.value,
    yates            = chisq.test(tab, correct = TRUE)$p.value,
    likelihood_ratio = oracle_lr(tab),
    n1_chisq         = oracle_n1(tab),
    fisher           = fisher.test(tab)$p.value,
    fisher_central   = oracle_fisher_central(tab),
    fisher_midp      = oracle_midp_central(tab),
    fisher_midp_sas  = oracle_midp_sas(tab),
    wald_z           = oracle_wald_z(tab),
    stop("unknown method")
  )), error = function(e) NA_real_)
  as.numeric(out)
}
```

## Functions

### Data-generating process

```{r}
generate_table <- function(n_per_group, pi_1, pi_2) {
  s1 <- rbinom(1, n_per_group, pi_1)
  s2 <- rbinom(1, n_per_group, pi_2)
  matrix(c(s1, n_per_group - s1,
           s2, n_per_group - s2),
         nrow = 2, byrow = TRUE)
}
```

### Analysis (compute the "reported" p under one analytic choice)

```{r}
generate_and_analyse <- function(n_per_group, pi_1, delta,
                                 analysis_method, p_digits) {
  pi_2 <- pi_1 + delta
  tab  <- generate_table(n_per_group, pi_1, pi_2)
  s1 <- tab[1, 1]; s2 <- tab[2, 1]

  p_true <- oracle_p(tab, analysis_method)
  if (is.na(p_true)) {
    return(tibble(s1, s2, n1 = n_per_group, n2 = n_per_group,
                  p_true = NA_real_, p_rep = NA_real_, p_operator = NA_character_))
  }

  p_rep <- roundwork::round_up(p_true, p_digits)
  if (dplyr::near(p_rep, 0)) {
    p_rep      <- 10^(-p_digits)
    p_operator <- "less_than"
  } else {
    p_operator <- "equals"
  }

  tibble(s1, s2, n1 = n_per_group, n2 = n_per_group, p_true, p_rep, p_operator)
}
```

### Reconstruction (evaluate inbounds for all three series)

```{r}
all_recalc_methods <- c("pearson", "yates", "likelihood_ratio", "n1_chisq",
                        "fisher", "fisher_central", "fisher_midp",
                        "fisher_midp_sas", "wald_z")

# Common rounded-bounds inbounds test (half-up rounding to p_digits), with
# censored-report handling matching the recalc convention.
inbounds_interval <- function(p_rep, p_operator, lo, hi, p_digits) {
  if (is.na(p_rep) || is.na(lo) || is.na(hi)) return(NA)
  lo_r <- roundwork::round_up(lo, p_digits)
  hi_r <- roundwork::round_up(hi, p_digits)
  if (p_operator == "equals")    return(p_rep >= lo_r & p_rep <= hi_r)
  if (p_operator == "less_than") return(lo_r <= p_rep)   # min bound below the < threshold
  NA
}

reconstruct_both <- function(reported, p_digits) {
  tab <- matrix(c(reported$s1, reported$n1 - reported$s1,
                  reported$s2, reported$n2 - reported$s2),
                nrow = 2, byrow = TRUE)

  # recalc full nine-method multiverse. Pass the reported p so `reproduced`
  # carries the DEFAULT (gap-aware union) decision and the union covered width.
  rc <- suppressMessages(suppressWarnings(recalc_chisq_p(
    counts     = tab,
    p          = reported$p_rep,
    p_digits   = p_digits,
    p_operator = reported$p_operator,
    p_methods  = all_recalc_methods,
    rounding   = "half_up"
  )))

  # reappraised span over its four tests
  rs <- reappraised_span(tab)

  # reappraised single default value
  rd <- tryCatch(reappraised_default(tab), error = function(e) NA_real_)

  tibble(
    # hull bounds (kept only for the interval-vs-p plot bands)
    min_p_recalc       = rc$reproduced$min_p,
    max_p_recalc       = rc$reproduced$max_p,
    # widths of the actual decision region: recalc's default is the UNION, so
    # its covered_width is the union region; hull_width kept for reference.
    covered_width_recalc = rc$reproduced$covered_width,
    hull_width_recalc    = rc$reproduced$hull_width,
    min_p_reapp        = rs[1],
    max_p_reapp        = rs[2],
    p_reapp_default    = rd,
    # recalc inbounds is the union (default); reappraised span/default are hulls/points
    inbounds_recalc    = rc$reproduced$p_inbounds,
    inbounds_reapp     = inbounds_interval(reported$p_rep, reported$p_operator, rs[1], rs[2], p_digits),
    inbounds_reapp_def = inbounds_interval(reported$p_rep, reported$p_operator, rd, rd, p_digits)
  )
}
```

## Define experiment

```{r}
experiment_parameters <- expand_grid(
  n_per_group     = c(20L, 80L, 200L),
  pi_1            = c(0.3, 0.5),
  delta           = c(0, 0.2, 0.4),
  analysis_method = all_recalc_methods,   # the analyst's true (unknown) choice
  p_digits        = c(2L, 3L),            # reporting-precision sweep
  iteration       = 1:50L # 50 for testing, e.g. 1:1500L for full sim
) |>
  mutate(condition = paste0(
    "n=",      n_per_group,
    " | pi1=", pi_1,
    " | delta=", delta,
    " | method=", analysis_method,
    " | pd=", p_digits
  ))
```

## Run simulation

```{r}
if (file.exists("simulation_p_recalc_vs_reappraised.rds")) {
  simulation <- read_rds("simulation_p_recalc_vs_reappraised.rds")
} else {

  set.seed(42)

  simulation <- experiment_parameters |>
    mutate(reported = future_pmap(
      .l = list(n_per_group     = n_per_group,
                pi_1            = pi_1,
                delta           = delta,
                analysis_method = analysis_method,
                p_digits        = p_digits),
      .f = generate_and_analyse,
      .progress = TRUE,
      .options  = furrr_options(seed = TRUE)
    )) |>
    unnest_wider(reported) |>
    filter(!is.na(p_true)) |>
    nest(reported = c(s1, s2, n1, n2, p_true, p_rep, p_operator)) |>
    mutate(rebuilt = future_pmap(
      .l = list(reported = reported, p_digits = p_digits),
      .f = reconstruct_both,
      .progress = TRUE,
      .options  = furrr_options(seed = TRUE)
    )) |>
    unnest_wider(reported) |>
    unnest_wider(rebuilt)

  write_rds(simulation, "simulation_p_recalc_vs_reappraised.rds", compress = "gz")
}
```

# Results

## Numerical fidelity on the four shared tests

Before any coverage analysis, confirm that the two tools agree on the tests they both implement. On a deterministic grid of 2×2 tables, `recalc`'s per-method *p*-value is compared against `reappraised`'s value for the corresponding `stat`. Differences should be at floating-point tolerance (the SAS mid-*p* carries `reappraised`'s own `round(., 9)`).

```{r}
shared_map <- tibble::tribble(
  ~recalc_method,     ~reappraised_stat,
  "pearson",          "chisq",
  "fisher",           "fisher",
  "fisher_midp",      "midp.epitools",
  "fisher_midp_sas",  "midp.sas"
)

fidelity_tables <- list(
  matrix(c(10,  2,  3, 15), 2, byrow = TRUE),
  matrix(c( 8,  7,  9,  6), 2, byrow = TRUE),
  matrix(c( 1,  9, 10,  2), 2, byrow = TRUE),
  matrix(c(20, 30, 25, 25), 2, byrow = TRUE),
  matrix(c( 0,  5,  6,  4), 2, byrow = TRUE),
  matrix(c(50, 50, 60, 40), 2, byrow = TRUE),
  matrix(c(15,  1,  2, 14), 2, byrow = TRUE)
)

recalc_method_p <- function(tab, method) {
  r <- suppressMessages(suppressWarnings(recalc_chisq_p(counts = tab, p_methods = method)))
  if (nrow(r$p_results) == 0) return(NA_real_)
  r$p_results$p_unrounded[r$p_results$method == method][1]
}

fidelity <- map_dfr(seq_along(fidelity_tables), function(i) {
  tab <- fidelity_tables[[i]]
  shared_map |>
    mutate(table = i,
           recalc_p     = vapply(recalc_method,    function(m) recalc_method_p(tab, m), numeric(1)),
           reappraised_p = vapply(reappraised_stat, function(s) reappraised_cat_p(tab, s), numeric(1)),
           abs_diff     = abs(recalc_p - reappraised_p))
})

fidelity |>
  group_by(recalc_method, reappraised_stat) |>
  summarise(max_abs_diff = max(abs_diff, na.rm = TRUE),
            n_tables     = dplyr::n(),
            .groups = "drop") |>
  kable(digits = 12, caption = "Max |recalc - reappraised| per shared test") |>
  kable_styling(full_width = FALSE)
```

## Coverage when the analyst's method is unknown

Each reported *p* was produced by one of the nine tests (`analysis_method`). The three series are: `recalc` (nine-method span), `reappraised (span)` (four-method span), and `reappraised (default)` (single `chi_midp` value).

### Overall coverage

```{r}
simulation |>
  summarise(
    K                      = n(),
    coverage_recalc        = mean(inbounds_recalc,    na.rm = TRUE),
    coverage_reapp_span    = mean(inbounds_reapp,     na.rm = TRUE),
    coverage_reapp_default = mean(inbounds_reapp_def, na.rm = TRUE)
  ) |>
  mutate(
    mcse_recalc        = sqrt(coverage_recalc        * (1 - coverage_recalc)        / K),
    mcse_reapp_span    = sqrt(coverage_reapp_span    * (1 - coverage_reapp_span)    / K),
    mcse_reapp_default = sqrt(coverage_reapp_default * (1 - coverage_reapp_default) / K)
  ) |>
  kable(digits = 4,
        caption = "Overall coverage across all nine analytic choices: recalc / reappraised (span) / reappraised (default)") |>
  kable_styling(full_width = FALSE)
```

### Coverage by analytic choice (the key table)

Each cell is `recalc / reappraised (span) / reappraised (default)`. The `yates` and `likelihood_ratio` rows are `reappraised`'s failure modes.

```{r}
coverage_by_method <- simulation |>
  group_by(analysis_method) |>
  summarise(
    K                      = n(),
    coverage_recalc        = mean(inbounds_recalc,    na.rm = TRUE),
    coverage_reapp_span    = mean(inbounds_reapp,     na.rm = TRUE),
    coverage_reapp_default = mean(inbounds_reapp_def, na.rm = TRUE),
    .groups = "drop"
  )

coverage_by_method |>
  mutate(cell = paste0(
    scales::number(coverage_recalc,        accuracy = 0.001), " / ",
    scales::number(coverage_reapp_span,    accuracy = 0.001), " / ",
    scales::number(coverage_reapp_default, accuracy = 0.001)
  )) |>
  select(analysis_method, K, cell) |>
  kable(caption = "Coverage by the test that produced the report: recalc / reappraised (span) / reappraised (default)") |>
  kable_styling(full_width = FALSE)
```

### Coverage by analytic choice × sample size

`reappraised`'s misses concentrate in the sparse (small-n) regime, where the nine tests diverge most.

```{r}
simulation |>
  group_by(analysis_method, n_per_group) |>
  summarise(
    coverage_recalc     = mean(inbounds_recalc, na.rm = TRUE),
    coverage_reapp_span = mean(inbounds_reapp,  na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(cell = paste0(
    scales::number(coverage_recalc,     accuracy = 0.001), " / ",
    scales::number(coverage_reapp_span, accuracy = 0.001)
  )) |>
  select(analysis_method, n_per_group, cell) |>
  pivot_wider(names_from = n_per_group, values_from = cell) |>
  kable(caption = "recalc / reappraised (span) coverage by analytic choice and n per group") |>
  kable_styling(full_width = FALSE) |>
  add_header_above(c(" " = 1, "n per group" = 3))
```

### Plot: coverage by analytic choice

```{r fig.height=5, fig.width=10}
coverage_by_method |>
  select(analysis_method, recalc = coverage_recalc,
         `reappraised (span)` = coverage_reapp_span,
         `reappraised (default)` = coverage_reapp_default) |>
  pivot_longer(-analysis_method, names_to = "tool", values_to = "coverage") |>
  mutate(tool = factor(tool, levels = c("recalc", "reappraised (span)", "reappraised (default)"))) |>
  ggplot(aes(x = analysis_method, y = coverage, fill = tool)) +
  geom_col(position = position_dodge()) +
  geom_hline(yintercept = 1, linetype = "dotted") +
  scale_fill_viridis_d(name = NULL) +
  ylim(0, 1) +
  labs(x = "Test that produced the report (analyst's unknown choice)",
       y = "Coverage",
       title = "Coverage when the analyst's categorical test is unknown",
       subtitle = "yates and likelihood_ratio are outside reappraised's method set") +
  theme_linedraw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 25, hjust = 1))
```

### Plot: reported p vs each tool's interval

```{r fig.height=18, fig.width=10}
plot_df <- simulation |>
  # The default is a single value, so give it a zero-width interval
  # (min = max = p_reapp_default). Tool tokens must have no internal underscore
  # so the names_pattern can split cleanly (hence "reappspan" / "reappdefault").
  mutate(
    min_p_reappspan    = min_p_reapp,      max_p_reappspan    = max_p_reapp,      inbounds_reappspan    = inbounds_reapp,
    min_p_reappdefault = p_reapp_default,  max_p_reappdefault = p_reapp_default,  inbounds_reappdefault = inbounds_reapp_def
  ) |>
  select(p_true, p_rep, analysis_method,
         min_p_recalc,       max_p_recalc,       inbounds_recalc,
         min_p_reappspan,    max_p_reappspan,    inbounds_reappspan,
         min_p_reappdefault, max_p_reappdefault, inbounds_reappdefault) |>
  pivot_longer(
    cols          = c(starts_with("min_p_"), starts_with("max_p_"), starts_with("inbounds_")),
    names_to      = c(".value", "tool"),
    names_pattern = "^(.*)_(recalc|reappspan|reappdefault)$"
  ) |>
  mutate(tool = factor(dplyr::recode(tool,
                              recalc       = "recalc (9 methods)",
                              reappspan    = "reappraised (span)",
                              reappdefault = "reappraised (default)"),
                       levels = c("recalc (9 methods)", "reappraised (span)",
                                  "reappraised (default)")))

ggplot(plot_df,
       aes(x = p_true, y = p_rep, colour = inbounds)) +
  geom_abline(slope = 1, linetype = "dotted") +
  geom_linerange(aes(ymin = min_p, ymax = max_p), alpha = 0.25) +
  geom_point(alpha = 0.4, size = 0.6) +
  scale_colour_viridis_d(begin = 0.2, end = 0.7, name = "p_inbounds") +
  facet_grid(analysis_method ~ tool) +
  labs(x = "True (oracle) p under the reporting test",
       y = "Reported (rounded) p",
       subtitle = "Vertical bars = each tool's [min_p, max_p] hull (recalc's decision is the union of method cells within its bar; reappraised (default) is a single value, so its bar is a point)") +
  coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  theme_linedraw() +
  theme(panel.grid.minor = element_blank(), legend.position = "bottom",
        strip.text.x = element_text(size = 7))
```

### Coverage by DGP

Coverage of each series collapsed over the DGP cells (sample size × baseline rate × effect), to show where the gap lives independent of the analyst's chosen test.

```{r}
simulation |>
  group_by(n_per_group, pi_1, delta) |>
  summarise(
    K                      = n(),
    coverage_recalc        = mean(inbounds_recalc,    na.rm = TRUE),
    coverage_reapp_span    = mean(inbounds_reapp,     na.rm = TRUE),
    coverage_reapp_default = mean(inbounds_reapp_def, na.rm = TRUE),
    .groups = "drop"
  ) |>
  kable(digits = 4,
        caption = "Coverage by DGP cell: recalc / reappraised (span) / reappraised (default).") |>
  kable_styling(full_width = FALSE)
```

## Pairwise disagreements

Per-replication, the three series can each flag (out-of-bounds) or cover (in-bounds) the reported *p*. The joint pattern shows which tool is responsible for each disagreement — the categorical analogue of the *t*-test comparison's joint-pattern table. `recalc` is expected to be uniformly equal-or-more-permissive than either `reappraised` reading under method uncertainty (it spans a superset of tests), so there should be essentially no "recalc flags / reappraised covers" rows.

```{r}
simulation |>
  mutate(
    disagree_recalc_vs_span    = inbounds_recalc != inbounds_reapp,
    disagree_recalc_vs_default = inbounds_recalc != inbounds_reapp_def,
    disagree_span_vs_default   = inbounds_reapp  != inbounds_reapp_def
  ) |>
  summarise(
    n_total                    = n(),
    recalc_vs_reapp_span       = sum(disagree_recalc_vs_span,    na.rm = TRUE),
    recalc_vs_reapp_default    = sum(disagree_recalc_vs_default, na.rm = TRUE),
    reapp_span_vs_default      = sum(disagree_span_vs_default,   na.rm = TRUE)
  ) |>
  kable(caption = "Number of replications where each pair of series produces different inbounds flags.") |>
  kable_styling(full_width = FALSE)
```

```{r}
simulation |>
  mutate(pattern = dplyr::case_when(
    inbounds_recalc &  inbounds_reapp &  inbounds_reapp_def  ~ "all three cover",
    inbounds_recalc &  inbounds_reapp & !inbounds_reapp_def  ~ "recalc + span cover; default flags",
    inbounds_recalc & !inbounds_reapp &  inbounds_reapp_def  ~ "recalc + default cover; span flags",
    inbounds_recalc & !inbounds_reapp & !inbounds_reapp_def  ~ "recalc only covers",
    !inbounds_recalc &  inbounds_reapp &  inbounds_reapp_def ~ "reappraised covers, recalc flags",
    !inbounds_recalc &  inbounds_reapp & !inbounds_reapp_def ~ "span only covers, recalc flags",
    !inbounds_recalc & !inbounds_reapp &  inbounds_reapp_def ~ "default only covers, recalc flags",
    !inbounds_recalc & !inbounds_reapp & !inbounds_reapp_def ~ "all three flag",
    TRUE ~ "other / NA"
  )) |>
  count(pattern) |>
  arrange(desc(n)) |>
  kable(caption = "Joint inbounds pattern across recalc / reappraised (span) / reappraised (default).") |>
  kable_styling(full_width = FALSE)
```

## Coverage / width trade-off

The width that matters is the **decision region** — the set of *p*-values each rule actually declares consistent. For `recalc` that is its **default union** (`covered_width`), the total width of the per-method cells, *not* the convex hull; the hull would be wide, but the gaps between methods are rejected, so they don't count. For `reappraised (span)` the decision region is its contiguous `[min, max]` (it has no gap-aware rule, so the span *is* its covered width); for `reappraised (default)` it is a single value (≈ 0). This table pairs coverage with these covered widths.

The result is that there is essentially **no trade-off**: `recalc`'s union covers every method *and* declares a smaller region than `reappraised (span)`, because the span fills in the gaps that the union excludes. (`recalc`'s hull width is shown for reference — it is wide, but it is not the decision region.)

```{r}
simulation |>
  mutate(width_reapp = max_p_reapp - min_p_reapp) |>   # span = its covered region
  group_by(n_per_group) |>
  summarise(
    K                      = n(),
    recalc        = mean(inbounds_recalc,    na.rm = TRUE),
    reapp_span    = mean(inbounds_reapp,     na.rm = TRUE),
    reapp_default = mean(inbounds_reapp_def, na.rm = TRUE),
    .groups                = "drop"
  ) |>
  kable(digits = 4,
        caption = "Coverage by sample size. recalc width = union covered_width (the default decision region); recalc hull shown for reference; reappraised span width = its contiguous interval; default is a single value (~0 width), coverage only. MCSE for mean width = sd(width)/sqrt(K).") |>
  kable_styling(full_width = FALSE)
```

```{r}
simulation |>
  mutate(width_reapp = max_p_reapp - min_p_reapp) |>   # span = its covered region
  group_by(n_per_group) |>
  summarise(
    K                      = n(),
    width_recalc_union = mean(covered_width_recalc),
    mcse_recalc_union    = sd(covered_width_recalc) / sqrt(K),
    width_recalc_hull      = mean(hull_width_recalc),
    mcse_recalc_hull      = sd(hull_width_recalc) / sqrt(K),
    width_reapp_span   = mean(width_reapp),
    mcse_reapp_span      = sd(width_reapp) / sqrt(K),
    .groups                = "drop"
  ) |>
  kable(digits = 4,
        caption = "Covered-region width by sample size. recalc width = union covered_width (the default decision region); recalc hull shown for reference; reappraised span width = its contiguous interval; default is a single value (~0 width), coverage only. MCSE for mean width = sd(width)/sqrt(K).") |>
  kable_styling(full_width = FALSE)
```

### Width and coverage by p-value bin

The covered region varies along the *p*-axis: `reappraised`'s span is widest at moderate *p* (where its tests spread out) and shrinks toward the boundaries, whereas `recalc`'s union covered width stays small everywhere (it is only the method cells). Pairing covered width with coverage per *p*-bin shows the union is both narrower and higher-coverage across the axis; the `p < .05` bins are the practically important ones for inconsistency detection.

```{r}
p_breaks <- c(0, 0.001, 0.01, 0.05, 0.10, 0.50, 1.0)
simulation |>
  mutate(
    width_reapp = max_p_reapp - min_p_reapp,
    p_bin       = cut(p_true, p_breaks, include.lowest = TRUE)
  ) |>
  group_by(p_bin) |>
  summarise(
    K                      = n(),
    coverage_recalc        = mean(inbounds_recalc, na.rm = TRUE),
    coverage_reapp         = mean(inbounds_reapp,  na.rm = TRUE),
    covered_w_recalc_union = mean(covered_width_recalc),
    covered_w_reapp_span   = mean(width_reapp),
    max_w_recalc_union     = max(covered_width_recalc),
    max_w_reapp_span       = max(width_reapp),
    .groups                = "drop"
  ) |>
  kable(digits = 4,
        caption = "Coverage and covered-region width by p-value bin: recalc (union) vs reappraised (span).") |>
  kable_styling(full_width = FALSE)
```

```{r fig.height=5, fig.width=9}
simulation |>
  mutate(width_reapp = max_p_reapp - min_p_reapp) |>
  select(p_true, `recalc (union)` = covered_width_recalc,
         `reappraised (span)` = width_reapp) |>
  pivot_longer(-p_true, names_to = "tool", values_to = "width") |>
  ggplot(aes(x = p_true, y = width, colour = tool)) +
  geom_point(alpha = 0.15, size = 0.6) +
  geom_smooth(se = FALSE) +
  scale_colour_viridis_d(begin = 0.2, end = 0.7, name = NULL) +
  labs(x = "True (oracle) p", y = "Covered-region width",
       title = "Covered-region width vs p-value: recalc (union) vs reappraised (span)") +
  theme_linedraw() +
  theme(panel.grid.minor = element_blank(), legend.position = "bottom")
```

## Subset analysis: replicates with true `p` < .05

The *p* < .05 regime is where most published, significant categorical results live, and where a spurious inconsistency flag (a legitimately-reported value falling outside the tool's interval) is most costly. The `delta = 0.4` DGP level was added to populate it. This subsection restricts coverage and width to `p_true < 0.05`.

```{r}
sig_subset <- simulation |> filter(p_true < 0.05)
n_subset   <- nrow(sig_subset)
n_total    <- nrow(simulation)
```

The subset contains `r n_subset` of the `r n_total` total replicates (`r scales::percent(n_subset / n_total, 0.1)`).

### Coverage on the p < .05 subset

Each cell is `recalc / reappraised (span) / reappraised (default)`.

```{r}
sig_subset |>
  group_by(analysis_method) |>
  summarise(
    K                      = n(),
    coverage_recalc        = mean(inbounds_recalc,    na.rm = TRUE),
    coverage_reapp_span    = mean(inbounds_reapp,     na.rm = TRUE),
    coverage_reapp_default = mean(inbounds_reapp_def, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(cell = paste0(
    scales::number(coverage_recalc,        accuracy = 0.001), " / ",
    scales::number(coverage_reapp_span,    accuracy = 0.001), " / ",
    scales::number(coverage_reapp_default, accuracy = 0.001)
  )) |>
  select(analysis_method, K, cell) |>
  kable(caption = "Coverage on the p < .05 subset: recalc / reappraised (span) / reappraised (default).") |>
  kable_styling(full_width = FALSE)
```

### Covered-region width on the p < .05 subset

```{r}
sig_subset |>
  mutate(width_reapp = max_p_reapp - min_p_reapp) |>
  group_by(n_per_group) |>
  summarise(
    K                      = n(),
    covered_w_recalc_union = mean(covered_width_recalc),
    mcse_w_recalc_union    = sd(covered_width_recalc) / sqrt(K),
    max_w_recalc_union     = max(covered_width_recalc),
    covered_w_reapp_span   = mean(width_reapp),
    mcse_w_reapp_span      = sd(width_reapp) / sqrt(K),
    max_w_reapp_span       = max(width_reapp),
    .groups                = "drop"
  ) |>
  kable(digits = 5,
        caption = "Covered-region width on the p < .05 subset, by sample size: recalc (union covered_width) vs reappraised (span). MCSE for mean width = sd(width)/sqrt(K).") |>
  kable_styling(full_width = FALSE)
```

## Coverage by reporting precision

The precision sweep. `p_digits` controls how coarsely the reported *p* (and each tool's interval bounds) are rounded for the inbounds comparison; a coarser `p_digits = 2` collapses more small *p*'s to a censored `p < .01` report. Coverage of each series is shown per precision.

```{r}
simulation |>
  group_by(p_digits) |>
  summarise(
    K                      = n(),
    coverage_recalc        = mean(inbounds_recalc,    na.rm = TRUE),
    coverage_reapp_span    = mean(inbounds_reapp,     na.rm = TRUE),
    coverage_reapp_default = mean(inbounds_reapp_def, na.rm = TRUE),
    .groups = "drop"
  ) |>
  kable(digits = 4,
        caption = "Coverage by reported p precision: recalc / reappraised (span) / reappraised (default).") |>
  kable_styling(full_width = FALSE)
```

## Failure-mode breakdown

The fraction of reports that each `reappraised` reading **misses**, for the five tests outside its set, by sample size. (`recalc` covers these by construction since all five are in its method set.)

```{r}
simulation |>
  filter(analysis_method %in% c("yates", "likelihood_ratio", "n1_chisq",
                                "fisher_central", "wald_z")) |>
  group_by(analysis_method) |>
  summarise(
    K               = n(),
    miss_reapp_span = mean(!inbounds_reapp,     na.rm = TRUE),
    miss_reapp_def  = mean(!inbounds_reapp_def, na.rm = TRUE),
    cover_recalc    = mean(inbounds_recalc,     na.rm = TRUE),
    .groups = "drop"
  ) |>
  kable(digits = 4,
        caption = "Miss rate of reappraised on its five failure modes (Yates, likelihood ratio, N-1, central Fisher, Wald z), with recalc coverage for reference.") |>
  kable_styling(full_width = FALSE)
```

## Capability comparison

This table is a *package-capability* map: which tests each tool can compute, and the source-package output each reproduces. All nine `recalc` methods are exercised in the quantitative analyses above.

```{r}
tibble::tribble(
  ~test,                                    ~recalc,  ~reappraised, ~reproduces_default_output_of,
  "Pearson chi-square (no correction)",     "yes",    "yes (chisq)", "Stata tabulate, SPSS Pearson row, SAS Chi-Square row, gtsummary >=1.3.6; = pooled 2-proportion z^2",
  "Yates continuity-corrected (2x2)",       "yes",    "no",          "R chisq.test()/prop.test() default, SPSS Continuity Correction row, tableone default, gtsummary <1.3.6",
  "Likelihood ratio (G^2)",                 "yes",    "no",          "SPSS / SAS Likelihood Ratio row; DescTools::GTest",
  "N-1 / Campbell-Richardson chi-square",   "yes",    "no",          "SPSS Linear-by-Linear Association (2x2); SAS PROC FREQ single-stratum Mantel-Haenszel ((N-1)/N x Pearson)",
  "Fisher's exact (two-sided, minlike)",    "yes",    "yes (fisher)","SPSS / SAS / R fisher.test default",
  "Fisher's exact (two-sided, central)",    "yes",    "no",          "exact2x2(tsmethod='central'); 'double the one-sided' (some SPSS/Prism)",
  "Central mid-p (epitools)",               "yes",    "yes (midp)",  "StatXact / OpenEpi / epitools",
  "SAS PROC FREQ mid-p",                    "yes",    "yes (midp.sas)", "SAS PROC FREQ exact mid-p",
  "Unpooled / Wald two-proportion z",       "yes",    "no",          "hand-rolled / some calculators; != Pearson",
  "Input rounding (proportions -> counts)", "yes",    "no",          "reconstruct integer tables from rounded %",
) |>
  kable(caption = "Test coverage: recalc vs reappraised, and the source-package output each test reproduces. Generator note: the *p* is usually stamped by a Table-1 generator whose default picks one of these tests (e.g. tableone -> Yates, gtsummary >=1.3.6 -> uncorrected Pearson with auto-switch to Fisher when any expected cell <5); recalc's multiverse already spans those test values, so attribution is about which row, not whether it is covered. Boschloo/Barnard (unconditional exact) and Blaker remain out of scope.") |>
  kable_styling(full_width = FALSE)
```

# Second simulation: rejection power for incorrect p-values

The first simulation measures **sensitivity** — does each tool's interval contain a *p* it *should*? Every reported *p* there is, by construction, reproducible, so it cannot measure **specificity**: does the tool correctly **reject** a *p* that *no* legitimate test produces? Specificity is exactly where the gap-aware union should win, because the hull-based rules accept the between-method gaps.

This second simulation probes specificity directly. For each simulated 2×2 table we sweep a grid of candidate reported *p*-values across the whole axis (`0.001` to `0.999` in steps of `0.001`, i.e. the reporting-precision grid at `p_digits = 3`) and record, for each tool, the fraction of candidates it **rejects** (declares not reproducible). A sharper rule rejects more of the axis — higher power to catch an incorrect, miscalculated, or fabricated *p* — provided it does not also reject the values that *are* reproducible.

**Reference set.** A categorical *p* is "legitimately reproducible" iff it matches some standard test on the table; the union of `recalc`'s nine methods *is* that set, so we take it as ground truth. Against it, `recalc`'s union rule is exact by construction (it accepts exactly the reproducible cells); the value here is quantifying how far the hull rules deviate — `recalc (hull)` and `reappraised (span)` accept gap values they should reject (low power), and `reappraised (span)`/`(default)` additionally reject reproducible Yates / likelihood-ratio / non-default values (low sensitivity).

```{r}
p_grid    <- seq(0.001, 0.999, by = 0.001)   # candidate reported p-values
pd_sweep  <- 3L
ulp_sweep <- 10^(-pd_sweep)
tol_sweep <- ulp_sweep * 0.5

rejection_for_table <- function(s1, s2, n1, n2,
                                p_grid, p_digits = 3, rounding = "half_up") {
  tab <- matrix(c(s1, n1 - s1, s2, n2 - s2), nrow = 2, byrow = TRUE)
  rc  <- suppressMessages(suppressWarnings(
    recalc_chisq_p(counts = tab, p_methods = all_recalc_methods, rounding = rounding)))
  mi  <- rc$method_intervals
  if (is.null(mi) || nrow(mi) == 0) return(NULL)

  # recalc union: accept iff candidate matches any per-method cell
  A <- outer(p_grid, mi$p_min_rounded - tol_sweep, `>=`) &
       outer(p_grid, mi$p_max_rounded + tol_sweep, `<=`)
  union_accept <- rowSums(A) > 0

  # recalc hull: accept iff candidate in [min_p_rounded, max_p_rounded]
  hull_accept <- p_grid >= rc$reproduced$min_p_rounded - tol_sweep &
                 p_grid <= rc$reproduced$max_p_rounded + tol_sweep

  # reappraised span: accept iff candidate in [min, max] over its 4 tests
  rs    <- reappraised_span(tab)
  rs_lo <- roundwork::round_up(rs[1], p_digits)
  rs_hi <- roundwork::round_up(rs[2], p_digits)
  span_accept <- p_grid >= rs_lo - tol_sweep & p_grid <= rs_hi + tol_sweep

  # reappraised default: accept iff candidate matches the single chi_midp value
  rd  <- roundwork::round_up(reappraised_default(tab), p_digits)
  def_accept <- abs(p_grid - rd) <= tol_sweep

  truly_incorrect <- !union_accept   # ground truth: not reproducible by any method
  saferate <- function(x) if (length(x) == 0) NA_real_ else mean(x)

  tibble::tibble(
    n_per_group = n1,
    # overall rejection rate across the swept axis
    rej_recalc_union = mean(!union_accept),
    rej_recalc_hull  = mean(!hull_accept),
    rej_reapp_span   = mean(!span_accept),
    rej_reapp_def    = mean(!def_accept),
    # power = correct rejection among truly-incorrect candidates
    pow_recalc_union = saferate((!union_accept)[truly_incorrect]),  # = 1 by construction
    pow_recalc_hull  = saferate((!hull_accept)[truly_incorrect]),
    pow_reapp_span   = saferate((!span_accept)[truly_incorrect]),
    pow_reapp_def    = saferate((!def_accept)[truly_incorrect]),
    # sensitivity = acceptance among truly-reproducible candidates
    sens_recalc_union = saferate(union_accept[union_accept]),       # = 1 by construction
    sens_recalc_hull  = saferate(hull_accept[union_accept]),
    sens_reapp_span   = saferate(span_accept[union_accept]),
    sens_reapp_def    = saferate(def_accept[union_accept])
  )
}
```

```{r}
if (file.exists("simulation_reject.rds")) {
  rej <- read_rds("simulation_reject.rds")
} else {
  set.seed(123)
  tables_all <- simulation |> distinct(s1, s2, n1, n2)
  tables     <- tables_all |> slice_sample(n = min(800L, nrow(tables_all)))
  rej <- tables |>
    mutate(out = future_pmap(
      .l = list(s1 = s1, s2 = s2, n1 = n1, n2 = n2),
      .f = function(s1, s2, n1, n2) rejection_for_table(s1, s2, n1, n2, p_grid),
      .options = furrr_options(seed = TRUE)
    )) |>
    select(out) |>
    tidyr::unnest(out)
  write_rds(rej, "simulation_reject.rds", compress = "gz")
}
```

## Overall rejection rate across the swept p-axis

Mean fraction of the `r length(p_grid)`-point candidate grid each tool rejects, averaged over tables. The two `recalc` rules keep full sensitivity (first simulation), so their rejection is "clean"; `reappraised (default)` rejects the most but at a large sensitivity cost (next table).

```{r}
rej |>
  summarise(
    `recalc (union, default)` = mean(rej_recalc_union),
    `recalc (hull)`           = mean(rej_recalc_hull),
    `reappraised (span)`      = mean(rej_reapp_span),
    `reappraised (default)`   = mean(rej_reapp_def)
  ) |>
  tidyr::pivot_longer(everything(), names_to = "tool", values_to = "mean_rejection_rate") |>
  kable(digits = 4,
        caption = "Mean proportion of candidate p-values rejected (higher = more power to flag an incorrect p).") |>
  kable_styling(full_width = FALSE)
```

### By p-value band

The same rejection metric, broken into significance-relevant *p*-value bands (subsets of the swept axis). The sweep is extended into the tail (`p` down to 1e-4) so the `p ≤ .001` band is populated. The plot of this breakdown is in the "Rejection power by p-value band" section below.

```{r}
band_breaks <- c(0, 0.001, 0.02, 0.05, 1.0)
band_labels <- c("p ≤ .001", ".001 < p ≤ .02",
                 ".02 < p ≤ .05", "p > .05")
p_grid_band <- sort(unique(c(seq(0.0001, 0.0009, by = 0.0001),
                             seq(0.001,  0.999,  by = 0.001))))
p_band <- cut(p_grid_band, band_breaks, include.lowest = TRUE, labels = band_labels)

rejection_by_band <- function(s1, s2, n1, n2, p_grid, p_band, p_digits = 3) {
  tab <- matrix(c(s1, n1 - s1, s2, n2 - s2), nrow = 2, byrow = TRUE)
  tol <- 0.5 * 10^(-p_digits)
  rc  <- suppressMessages(suppressWarnings(
    recalc_chisq_p(counts = tab, p_methods = all_recalc_methods, rounding = "half_up")))
  mi  <- rc$method_intervals
  if (is.null(mi) || nrow(mi) == 0) return(NULL)
  union_acc <- rowSums(outer(p_grid, mi$p_min_rounded - tol, `>=`) &
                       outer(p_grid, mi$p_max_rounded + tol, `<=`)) > 0
  hull_acc  <- p_grid >= rc$reproduced$min_p_rounded - tol &
               p_grid <= rc$reproduced$max_p_rounded + tol
  rs <- reappraised_span(tab)
  span_acc <- p_grid >= roundwork::round_up(rs[1], p_digits) - tol &
              p_grid <= roundwork::round_up(rs[2], p_digits) + tol
  rd <- roundwork::round_up(reappraised_default(tab), p_digits)
  def_acc <- abs(p_grid - rd) <= tol
  rate <- function(acc) as.numeric(tapply(!acc, p_band, mean))  # rejection rate / band
  data.frame(band = levels(p_band),
             `recalc (union)`         = rate(union_acc),
             `recalc (hull)`          = rate(hull_acc),
             `reappraised (span)`     = rate(span_acc),
             `reappraised (default)`  = rate(def_acc),
             check.names = FALSE)
}

if (file.exists("simulation_reject_band.rds")) {
  band_rej <- read_rds("simulation_reject_band.rds")
} else {
  set.seed(123)  # same sample as the rejection run above
  tables_all <- simulation |> distinct(s1, s2, n1, n2)
  tables_b   <- tables_all |> slice_sample(n = min(800L, nrow(tables_all)))
  band_rej <- tables_b |>
    mutate(out = future_pmap(
      list(s1 = s1, s2 = s2, n1 = n1, n2 = n2),
      function(s1, s2, n1, n2) rejection_by_band(s1, s2, n1, n2, p_grid_band, p_band),
      .options = furrr_options(seed = TRUE)
    )) |>
    select(out) |>
    tidyr::unnest(out)
  write_rds(band_rej, "simulation_reject_band.rds", compress = "gz")
}

band_summary <- band_rej |>
  group_by(band) |>
  summarise(across(c(`recalc (union)`, `recalc (hull)`,
                     `reappraised (span)`, `reappraised (default)`), mean),
            .groups = "drop") |>
  mutate(band = factor(band, levels = band_labels)) |>
  arrange(band)

band_summary |>
  kable(digits = 4,
        caption = "Mean rejection rate within each p-value band, by rule. recalc's union rejects more than the hull rules (recalc hull, reappraised span) in the larger-p bands, where between-method gaps are widest; the rules converge in the smallest-p band as the methods cluster near 0. reappraised's single default rejects most everywhere, but that is over-rejection (it accepts only a fraction of genuinely reproducible values; see the power-vs-sensitivity table).") |>
  kable_styling(full_width = FALSE)
```

## Power vs sensitivity (against the recalc-union reference)

`power` = fraction of truly-incorrect candidates correctly rejected; `sensitivity` = fraction of truly-reproducible candidates correctly accepted. `recalc (union)` is 1 / 1 by construction; the hull rules trade one off against the other.

```{r}
rej |>
  summarise(
    `recalc (union)_power`    = mean(pow_recalc_union, na.rm = TRUE),
    `recalc (union)_sens`     = mean(sens_recalc_union, na.rm = TRUE),
    `recalc (hull)_power`     = mean(pow_recalc_hull,  na.rm = TRUE),
    `recalc (hull)_sens`      = mean(sens_recalc_hull, na.rm = TRUE),
    `reappraised (span)_power`= mean(pow_reapp_span,   na.rm = TRUE),
    `reappraised (span)_sens` = mean(sens_reapp_span,  na.rm = TRUE),
    `reappraised (default)_power` = mean(pow_reapp_def, na.rm = TRUE),
    `reappraised (default)_sens`  = mean(sens_reapp_def, na.rm = TRUE)
  ) |>
  tidyr::pivot_longer(everything(),
    names_to = c("tool", ".value"), names_sep = "_(?=power$|sens$)") |>
  kable(digits = 4,
        caption = "Power (correct rejection of incorrect p) and sensitivity (acceptance of reproducible p). recalc union dominates: power AND sensitivity = 1.") |>
  kable_styling(full_width = FALSE)
```

## Rejection power by sample size

The union's advantage is largest for sparse (small-*n*) tables, where the methods spread out and the hull contains the widest gaps.

```{r}
rej |>
  group_by(n_per_group) |>
  summarise(
    K                      = n(),
    `recalc (union)`       = mean(rej_recalc_union),
    `recalc (hull)`        = mean(rej_recalc_hull),
    `reappraised (span)`   = mean(rej_reapp_span),
    .groups = "drop"
  ) |>
  kable(digits = 4,
        caption = "Mean rejection rate by sample size: recalc union rejects more of the axis than the hull rules, most so at small n.") |>
  kable_styling(full_width = FALSE)
```

## Rejection power by p-value band

Where on the *p*-axis does each rule's rejection power lie? This bins the swept
candidate *p* into significance-relevant bands and reports the rejection rate of
each rule within each band. The sweep is extended into the tail
(`p` down to 1e-4) so the `p ≤ .001` band is populated. The union's gap-aware
advantage over the hull rules is largest in the **larger-*p* bands**, where the
methods spread out and a contiguous hull swallows wide between-method gaps; in
the smallest-*p* band the methods cluster near zero, so all interval rules
behave similarly. The practically important point is that the union stays
high-rejection across the small-*p* bands too — catching an incorrect
significant *p* — *without* the sensitivity loss that gives `reappraised`'s
single default its (over-)high rejection. The per-band rejection rates are
tabulated in the *Overall rejection rate* section above (`band_summary`); the
chart below visualises them.

```{r fig.height=4.5, fig.width=9}
band_summary |>
  tidyr::pivot_longer(-band, names_to = "rule", values_to = "rejection_rate") |>
  mutate(rule = factor(rule, levels = c("recalc (union)", "recalc (hull)",
                                        "reappraised (span)", "reappraised (default)"))) |>
  ggplot(aes(x = band, y = rejection_rate, fill = rule)) +
  geom_col(position = position_dodge()) +
  scale_fill_viridis_d(name = NULL) +
  ylim(0, 1) +
  labs(x = "Candidate reported-p band", y = "Rejection rate",
       title = "Rejection power by p-value band",
       subtitle = "recalc's union beats the hull rules most in the larger-p bands; reappraised's default rejects most but over-rejects (low sensitivity)") +
  theme_linedraw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 15, hjust = 1))
```

## Illustration: accept/reject across the p-axis for one table

For the worked table, the candidates each tool **accepts** (coloured) against the full axis. `recalc (union)` accepts only the handful of method cells; `recalc (hull)` and `reappraised (span)` accept whole contiguous bands (including the gaps); `reappraised (default)` accepts a single cell. Everything not coloured is rejected.

```{r fig.height=3.2, fig.width=9}
ex <- c(s1 = 12, n1 = 20, s2 = 8, n2 = 20)
ex_tab <- matrix(c(ex["s1"], ex["n1"] - ex["s1"],
                   ex["s2"], ex["n2"] - ex["s2"]), nrow = 2, byrow = TRUE)
ex_grid <- seq(0.001, 0.6, by = 0.001)
rc_ex   <- recalc_chisq_p(counts = ex_tab, p_methods = all_recalc_methods, rounding = "half_up")
mi_ex   <- rc_ex$method_intervals
rs_ex   <- reappraised_span(ex_tab)
rd_ex   <- roundwork::round_up(reappraised_default(ex_tab), 3)

acc <- function(name, accepted) tibble::tibble(tool = name, p = ex_grid, accepted = accepted)
ex_df <- dplyr::bind_rows(
  acc("recalc (union)", rowSums(outer(ex_grid, mi_ex$p_min_rounded - tol_sweep, `>=`) &
                                 outer(ex_grid, mi_ex$p_max_rounded + tol_sweep, `<=`)) > 0),
  acc("recalc (hull)", ex_grid >= rc_ex$reproduced$min_p_rounded - tol_sweep &
                       ex_grid <= rc_ex$reproduced$max_p_rounded + tol_sweep),
  acc("reappraised (span)", ex_grid >= roundwork::round_up(rs_ex[1], 3) - tol_sweep &
                            ex_grid <= roundwork::round_up(rs_ex[2], 3) + tol_sweep),
  acc("reappraised (default)", abs(ex_grid - rd_ex) <= tol_sweep)
) |>
  mutate(tool = factor(tool, levels = c("recalc (union)", "recalc (hull)",
                                        "reappraised (span)", "reappraised (default)")))

ggplot(ex_df, aes(x = p, y = tool, fill = accepted)) +
  geom_tile(height = 0.7) +
  scale_fill_manual(values = c(`TRUE` = "#21908C", `FALSE` = "grey92"),
                    labels = c(`TRUE` = "accepted (consistent)", `FALSE` = "rejected"),
                    name = NULL) +
  labs(x = "Candidate reported p", y = NULL,
       title = "What each rule accepts across the p-axis (table: 12/8 vs 8/12)",
       subtitle = "recalc union accepts only the method cells; the hull rules accept the gaps too") +
  theme_linedraw() +
  theme(panel.grid = element_blank(), legend.position = "bottom")
```

# Discussion

<!-- Discuss and interpret results of the simulation here. -->

## Numerical fidelity

On the four tests both tools implement, `recalc` reproduces `reappraised`'s value to floating-point tolerance: Pearson, Fisher, and the central mid-*p* at machine epsilon, and the SAS mid-*p* at the ~1e-9 level carried by `reappraised`'s own `round(., 9)`. The coverage differences below therefore come entirely from method *breadth*, not from disagreement on shared methods.

## Coverage when the analyst's method is unknown

Treating each tool as a multiverse over the methods it knows, `recalc` covers the reported *p* for **every** analytic choice, because all nine candidate tests are in its span. `reappraised (span)` covers the four tests it implements but **misses** a meaningful fraction of the five reports it cannot reproduce (`yates`, `likelihood_ratio`, `n1_chisq`, `fisher_central`, `wald_z`) — including R's and SPSS/SAS's default output (Yates, likelihood ratio, the SPSS Linear-by-Linear / SAS Mantel-Haenszel N−1 value). The miss rate is largest in the sparse small-*n* cells, where the tests diverge most, and shrinks toward 0 as *n* grows and they converge.

`reappraised (default)` — the single `chi_midp` value the package emits out of the box — covers only the analytic choices whose *p* happens to round to that one number, so its coverage is high only for `pearson`-style reports (and `midp` in the sparse cells where the fallback fires) and low elsewhere. This is the realistic cost of a point reproduction when the analyst's method is unknown.

## Coverage / width trade-off

There is essentially **no** coverage/width trade-off once the right width is used. `recalc`'s *hull* is wide — its nine methods include the Yates and central-Fisher values (near or above Fisher) and the likelihood-ratio and Wald-z values (below Pearson) — but the hull is not the decision region. The default rule is the gap-aware **union**, whose covered width is just the handful of method cells, and that is typically *smaller* than `reappraised`'s contiguous four-test span (which fills in the between-method gaps). So `recalc` covers every method **and** declares a smaller consistent region than `reappraised (span)` — it dominates on both axes rather than trading one for the other. The covered-width tables use this union width (the hull is reported only for reference); the rejection-power simulation below quantifies the same advantage by sweeping the *p*-axis.

## Rejection power (second simulation)

The width result above is about the *hull*; the rejection result is about what the hull conceals. Sweeping the whole *p*-axis shows that `recalc`'s default **union** rule rejects far more of the axis than either hull rule, because it accepts only the discrete method cells and rejects the between-method gaps — while still accepting every reproducible value (sensitivity = 1). `recalc (hull)` and `reappraised (span)` accept those gaps, so they have lower power to flag an incorrect *p*; the advantage is largest for sparse tables, where the gaps are widest. `reappraised (default)` rejects the most of all, but that is over-rejection: its sensitivity is a small fraction (it accepts only its single `chi_midp` value and rejects the reproducible values from every other test). The upshot is that the union rule is the only one that is simultaneously high-power and high-sensitivity — it dominates the hull rules on the specificity axis at no cost on the sensitivity axis. The headline first-simulation coverage numbers (all ≈ 1) hide this entirely, which is why the second simulation is needed.

## Conclusions

1. **Fidelity** — `recalc` reproduces all four of `reappraised`'s categorical tests to floating-point tolerance.
2. **Coverage under method uncertainty** — `recalc` covers all nine analytic choices; `reappraised` misses a quantified fraction of the five reports outside its set (its span) or most non-default reports (its single default value).
3. **Trade-off** — `recalc`'s broader coverage costs interval width, concentrated in the sparse regime; the two converge as *n* grows.

As with the continuous comparison, this says nothing about `reappraised`'s *primary* claim — the distribution of baseline categorical *p*-values across a corpus of trials evaluated against a reference distribution (the AUC machinery of `pval_cat_fn()`). That use case is upstream of the per-table *p*-value computation exercised here.

## Limitations of this simulation

- **2×2 tables only.** Both tools handle larger tables for the asymptotic and Fisher tests; that path is not compared here.
- **Two independent binomials.** A fixed-total or case-control design would change the sampling distribution but not the tests' *p*-value definitions.
- **`reappraised`'s machinery is exercised outside its usual entry point.** Its per-table code is lifted out of `pval_cat_fn()`; the reference-distribution / AUC steps are not run.
- **This benchmarks the `reappraised` package, not its Shiny app.** The `reappraised` Shiny app (`apps/app_dichotomous.R`) is a separate, more capable tool: it computes seven tests (Pearson, Yates, likelihood ratio, CMH/Mantel-Haenszel = N−1, Fisher minlike, epitools mid-*p*, SAS mid-*p*) and flags whether the reported *p* matches **any** of them — a gap-aware *union* like `recalc` (covering seven of `recalc`'s nine methods; it lacks only `fisher_central` and `wald_z`), plus a missing-data reconstruction search. So the span/default results here understate what the app does; the app and `recalc` are close on the method-breadth axis.
- **Version-specific.** Results are pinned to `reappraised` v0.2 (CRAN, 2026-03-05); a future version that adds Yates or likelihood-ratio options would change the coverage gap.
- **Precision sweep is coarse-grained.** Only `p_digits` ∈ {2, 3} is swept (cell counts are exact integers, so there is no proportion-precision axis on the categorical side); finer or coarser *p* reporting could be added.
- **Specificity is assessed by sweep, not by a fitted error model.** The second simulation measures specificity by sweeping the whole *p*-axis and asking what fraction each rule rejects, taking `recalc`'s nine-method union as the reference set of reproducible values. This is a deliberately simple error model (every off-reference value is treated as equally likely to be reported in error); a more realistic error-injection DGP — perturbing a true *p* by plausible typo/miscalculation mechanisms, or computing it from a slightly different table — would weight the *p*-axis by how errors actually arise and is left to future work.

# References

Bolland, M. J., Gamble, G. D., Avenell, A., Grey, A., & Lumley, T. (2019). Baseline P value distributions in randomized trials were uniform for continuous but not categorical variables. *Journal of Clinical Epidemiology*, 112, 67–76.

Bolland, M. (2024). *reappraised: Statistical tools for assessing publication integrity of groups of trials* [R package, version 0.2]. CRAN: https://CRAN.R-project.org/package=reappraised.

Morris, T. P., White, I. R., & Crowther, M. J. (2019). Using simulation studies to evaluate statistical methods. *Statistics in Medicine*, 38(11), 2074–2102.

Siepe, B. S., Bartoš, F., Morris, T. P., Boulesteix, A.-L., Heck, D. W., & Pawel, S. (2024). Simulation studies for methodological research in psychology: A standardized template for planning, preregistration, and reporting. *Psychological Methods*.

# Session info

```{r}
sessionInfo()
```
